### 2024-11-26 paper/paper.qmd
Set up, make a title and introduction

---
title: "Canada’s Battle Against Car Theft: Progress Amid Persistent Challenges "
subtitle: "Government measures and public vigilance show promise, but organized theft networks remain a concern."
author: 
  - First author
  - Another author
thanks: "Code and data are available at: [https://github.com/RohanAlexander/starter_folder](https://github.com/RohanAlexander/starter_folder)."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(palmerpenguins)
```

# Introduction

Car theft has long been a concern in Canada, but recent years have witnessed a dramatic surge in reported cases, driven by sophisticated theft operations. Criminal networks often target vehicles with advanced keyless entry systems, exploiting vulnerabilities in wireless technology. These stolen vehicles are frequently exported overseas or dismantled for parts, making recovery difficult.

In 2024, the Insurance Bureau of Canada reported a 19% reduction in theft claims during the first half of the year, attributed to increased law enforcement efforts and public awareness campaigns. Notable initiatives include the Canadian government’s National Action Plan to Combat Car Theft, which emphasizes collaboration between federal agencies, law enforcement, and border security.

Despite these efforts, challenges persist. Border authorities continue to intercept stolen vehicles destined for international markets, and data suggests that many thefts remain unreported or unresolved. The fight against car theft in Canada highlights the need for continuous innovation, community cooperation, and robust legal frameworks to address this evolving crime. This article delves into the current state of car theft in Canada, recent government actions, and the ongoing challenges in combating this widespread crime.

Estimand paragraph

Results paragraph

Why it matters paragraph

Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....

# Data {#sec-data}

## Overview

We use the statistical programming language R [@citeR].... Our data [@shelter].... Following @tellingstories, we consider...

Overview text

## Measurement

Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.

## Outcome variables

Add graphs, tables and text. Use sub-sub-headings for each outcome variable or update the subheading to be singular.

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false

ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = "none") +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false

analysis_data <- read_csv(here::here("data/02-analysis_data/analysis_data.csv"))

analysis_data |> 
  ggplot(aes(x = width, y = length)) +
  geom_point(alpha = 0.8) +
  theme_minimal() +
  labs(x = "Wing width (mm)",
       y = "Wing length (mm)")
```

Talk way more about it.

## Predictor variables

Add graphs, tables and text.

Use sub-sub-headings for each outcome variable and feel free to combine a few into one if they go together naturally.

# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

first_model <-
  readRDS(file = here::here("models/first_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

modelsummary::modelsummary(
  list(
    "First model" = first_model
  ),
  statistic = "mad",
  fmt = 2
)
```

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

Please don't use these as sub-heading labels - change them to be what your point actually is.

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(first_model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(first_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(first_model, "trace")

plot(first_model, "rhat")
```

\newpage

# References

### 2024-11-26 paper/paper.qmd
minorchange1.3
minorchange1.x is used to update title, structure and introduction.

@@
 author: 
-  - First author
-  - Another author
-thanks: "Code and data are available at: [https://github.com/RohanAlexander/starter_folder](https://github.com/RohanAlexander/starter_folder)."
+  - Yiyue Deng
+thanks: "Code and data are available at: [https://github.com/vanessadyy/Car_Theft](https://github.com/vanessadyy/Car_Theft)."
 date: today

### 2024-12-01 paper/paper.qmd
MethodUpdate
Data and method update

@@
 abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
-format: pdf
+include-before:
+- '\newpage'
+format:
+  pdf:
+    toc: true
+    toc-depth: 2
     number-sections: true
 bibliography: references.bib
+csl: apa6.csl
+editor_options: 
+chunk_output_type: console
 ---
@@
 library(tidyverse)
-library(palmerpenguins)
+library(tidyverse)
+library(here)
+library(lubridate)
+library(ggplot2)
+library(scales)
+library(knitr)
+library(kableExtra)
 ```
+
+\newpage
 
@@
 
-Car theft has long been a concern in Canada, but recent years have witnessed a dramatic surge in reported cases, driven by sophisticated theft operations.
-Criminal networks often target vehicles with advanced keyless entry systems, exploiting vulnerabilities in wireless technology.
-These stolen vehicles are frequently exported overseas or dismantled for parts, making recovery difficult.
+Vehicle theft has become a widespread public concern in Canada, leaving many citizens feeling helpless and frustrated.
+As car theft incidents continue to rise, police chiefs have made surprising recommendations, including advising residents to leave their car keys near the door to avoid violent confrontations during break-ins.
+This suggestion sparked public outrage, with many criticizing it online as absurd and unreasonable.
+The widespread public discourse escalated the issue into a national battle against crime.
 
-In 2024, the Insurance Bureau of Canada reported a 19% reduction in theft claims during the first half of the year, attributed to increased law enforcement efforts and public awareness campaigns.
-Notable initiatives include the Canadian government’s National Action Plan to Combat Car Theft, which emphasizes collaboration between federal agencies, law enforcement, and border security.
+## The Scale of the Problem
 
-Despite these efforts, challenges persist.
-Border authorities continue to intercept stolen vehicles destined for international markets, and data suggests that many thefts remain unreported or unresolved.
-The fight against car theft in Canada highlights the need for continuous innovation, community cooperation, and robust legal frameworks to address this evolving crime.
-This article delves into the current state of car theft in Canada, recent government actions, and the ongoing challenges in combating this widespread crime.
+The scale of vehicle theft in Canada is alarming, exposing government inefficiency in tackling the problem.
+The reaction from law enforcement has been equally unexpected, with officers treating theft as an inevitable occurrence and even providing rationales for its prevalence.
+While insurance companies offer quick resolutions for stolen vehicle claims, this comes at a cost.
+Insurers are not charitable organizations, and frequent claims could lead to higher premiums or even the loss of insurance eligibility for some car owners.
+The reliance on insurance as a remedy highlights the authorities' decision to abandon pursuit of stolen vehicles, shifting the burden onto vehicle owners to seek reimbursement.
+
+## The Vicious Cycle of Vehicle Theft
+
+The indifferent attitude of law enforcement, coupled with lenient legal consequences, has fueled an annual increase in car thefts, pushing Canada into a vicious cycle of buying and stealing vehicles.
+The legal responsibilities surrounding car theft and the evasive accountability of international import companies reflect the complexities of organized crime.
+A large portion of stolen vehicles ends up in the Middle East and Africa, particularly Nigeria, where demand for high-quality second-hand cars is strong.
+The COVID-19 pandemic and geopolitical conflicts have exacerbated supply shortages, causing car prices in the Middle East to surge and further incentivizing theft.
+
+## The Global Supply Chain of Stolen Vehicles
+
+The vehicle theft industry operates as an international supply chain, where cars of varying quality are allocated to different markets, creating a clear hierarchy.
+High-value vehicles are often shipped to affluent buyers, while lower-tier models serve other regions.
+The organized and systematic nature of these operations underscores the challenges faced by Canadian authorities in addressing this growing issue.
+Without comprehensive legal reforms and stricter enforcement measures, the country risks further entrenchment in this illicit cycle, with profound economic and social consequences.
+
+## Public Response and Anti-Theft Measures
+
+Faced with insufficient responses from the police, many vehicle owners have begun implementing their own anti-theft measures.
+However, professional thieves often view these devices as minor obstacles, and even robust anti-theft systems have limitations.
+High installation costs and vulnerabilities in extreme weather conditions further undermine their effectiveness.
+Some individuals attempt to protect their vehicles by parking in neighborhoods with higher police presence or near police stations.
+However, such measures do not guarantee safety, as evidenced by reports of police chiefs’ vehicles being stolen in broad daylight.
+These incidents highlight the audacity of thieves and have led to online ridicule of police authorities for failing to practice basic precautions.
+
+## The National Action Plan
+
+Under pressure from the public and media, the Canadian government introduced a series of policies in early 2024 as part of a “National Action Plan” to combat auto theft.
+The key components of the plan include:
+
+- **Amendments to Criminal Laws**: Revising legislation to provide law enforcement and prosecutors with enhanced tools to tackle vehicle theft.
+- **Changes to the Radiocommunication Act**: Tightening regulations on radio devices commonly used in car thefts.
+- **Enhanced Intelligence and Information Sharing**: Promoting better collaboration among municipal, provincial, federal, and international agencies.
+- **Technology Deployment**: Leveraging advanced technologies such as scanning systems, data analytics, and GPS tracking to enhance inspection efficiency, particularly at ports and rail terminals.
+- **Specialized Training**: Offering targeted training programs for law enforcement officers to improve their capacity to investigate vehicle theft cases.
+- **Intergovernmental Task Force**: Creating a National Vehicle Theft Task Force to enhance coordination among government bodies.
+- **Support for Anti-Theft Technology Development**: Encouraging innovation and wider adoption of commercial anti-theft technologies.
+
+## Interim Result and Aim of the Paper
+
+Several months have passed since the plan's implementation.
+Open Data Toronto has published theft-from-motor-vehicle data for Toronto reported from January 2014 to September 2024.
+To assess interim results, we analyzed this dataset to evaluate the effectiveness of the government's actions.
+Our objective is to produce statistically objective conclusions that quantify the plan's impact.
+
+# Data {#sec-data}
+
+Since 2014, the Toronto Police Service has been collecting auto theft data in digital format.
+The data is updated and maintained quarterly, with public access provided via Open Data Toronto.
+This dataset includes approximate locations of car theft incidents, facilitating the analysis of geographic trends over time.
+The latest update, released on October 19, 2024, covers reported theft cases up to September 30, 2024.
+
+## Data Source
+
+The dataset used in this study was obtained from [Open Data Toronto](https://open.toronto.ca/dataset/deaths-of-shelter-residents/).
+The raw dataset comprises the following [key fields](@tbl-variable_description):
+
+```{r}
+#| label: tbl-variable_description
+#| tbl-cap: "Raw Variable Description"
+#| echo: false
+#| warning: false
+#| message: false
+
+variable_description <- data.frame(
+  Column = c(
+    "_id", "EVENT_UNIQUE_ID", "REPORT_DATE", "OCC_DATE", "REPORT_YEAR",
+    "REPORT_MONTH", "REPORT_DAY", "REPORT_DOY", "REPORT_DOW", "REPORT_HOUR",
+    "OCC_YEAR", "OCC_MONTH", "OCC_DAY", "OCC_DOY", "OCC_DOW", "OCC_HOUR",
+    "DIVISION", "LOCATION_TYPE", "PREMISES_TYPE", "UCR_CODE", "UCR_EXT",
+    "OFFENCE", "MCI_CATEGORY", "HOOD_158", "LONG_WGS84", "LAT_WGS84"
+  ),
+  Description = c(
+    "Unique row identifier for Open Data database",
+    "Offence Number",
+    "Date Offence was Reported",
+    "Date of Offence",
+    "Year Offence was Reported",
+    "Month Offence was Reported",
+    "Day of the Month Offence was Reported",
+    "Day of the Year Offence was Reported",
+    "Day of the Week Offence was Reported",
+    "Hour Offence was Reported",
+    "Year Offence Occurred",
+    "Month Offence Occurred",
+    "Day of the Month Offence Occurred",
+    "Day of the Year Offence Occurred",
+    "Day of the Week Offence Occurred",
+    "Hour Offence Occurred",
+    "Police Division where Offence Occurred",
+    "Location Type of Offence",
+    "Premises Type of Offence",
+    "UCR Code for Offence",
+    "UCR Extension for Offence",
+    "Title of Offence",
+    "MCI Category of Occurrence",
+    "Identifier of Neighbourhood using City of Toronto's new 158 neighbourhood structure",
+    "Longitude coordinate",
+    "Latitude coordinate"
+  )
+)
+# Render the table
+kable(variable_description, col.names = c("Column", "Description")) %>% kable_styling(full_width = FALSE) %>%  column_spec(1, width = "4cm") %>% column_spec(2, width = "8cm")
+```
+
+## Data Notes
+
+This dataset provides detailed information about auto theft cases reported in Toronto.
+Each row corresponds to a unique occurrence and includes both temporal and spatial details about the offence.
+The data can be used to analyze trends over time and across neighborhoods in the city.
+
+- **Temporal Variables**: Columns such as `REPORT_DATE`, `REPORT_YEAR`, and `OCC_DATE` provide insights into when offences were reported or occurred.
+
+- **Spatial Variables**: Columns like `DIVISION`, `HOOD_158`, `LONG_WGS84`, and `LAT_WGS84` enable spatial analysis and visualization of offences across Toronto.
+
+- **Categorical Variables**: Fields such as `LOCATION_TYPE`, `PREMISES_TYPE`, and `OFFENCE` classify the nature and context of the reported incidents.
+
+## Variables of Interest
+
+Dependent or duplicated information was recorded using different variables in the dataset.
+To avoid redundancy, for each scenario, we selected representative variables from the categories of temporal, spatial, and crime severity information related to theft crimes.
+The selected variables are outlined below:
+
+- **Temporal Information**
+We selected variables representing the occurrence date and excluded the reported date. The following variables were retained for analysis:
+  - OCC_DATE
+  - OCC_YEAR
+  - OCC_MONTH
+  - OCC_DAY
+  - OCC_DOW
+  - OCC_HOUR
+
+- **Spatial Information**
+Key spatial information is represented by the police division number, as well as the longitude and latitude of the theft location.
+Additionally, the location type was included for further context.
+The following variables were retained:
+  - DIVISION
+  - LONG_WGS84
+  - LAT_WGS84
+  - PREMISES_TYPE
+
+- **Crime Severity Information**
+To capture the severity of theft crimes, we selected the UCR code and offense title as indicators. The following variables were retained:
+  - UCR_CODE
+  - OFFENCE
+
+## Software and Packages
+
+We used R 4.3 for analysis and followed a public template @citeR@citeRohan. and a collection of R packages were used in this study:
+
+- tidyverse @Citetidyverse
+- here @Citehere
+- lubridate @Citelubridate
+- ggplot2 @Citeggplot2
+- scales @Citescales
+- knitr @Citeknitr
+- kableExtra
+
+## Data Management
+
+Data cleaning steps are described in the [Appendix](#sec-appendix).
+A sample of the cleaned data is presented in Table 2-4.
+Among all variables, Year, Total_Death, Male, and Female are coded as integers, Month is coded as abbreviations in character format, and Time is coded in date format.
+After data cleaning, there are no missing values remaining in the dataset.
+
+```{r}
+#| echo: false
+#| warning: false
+
+i_am("README.md")
+dt <- read_csv(here("data","analysis_data","analysis_data.csv"))
+data_part1 <- kable(head(dt)[, 1:6],caption = "Sample of Clean Data: Part 1")
+data_part2 <- kable(head(dt)[, 7:10],caption = "Sample of Clean Data: Part 2")
+data_part3 <- kable(head(dt)[, 11:13],caption = "Sample of Clean Data: Part 3")
+data_part1
+data_part2
+data_part3
+```
 
@@
 
-We use the statistical programming language R [@citeR].... Our data [@shelter].... Following @tellingstories, we consider...
+We use the statistical programming language R .... Our data .... Following , we consider...
 
@@
 
-Some of our data is of penguins (@fig-bills), from @palmerpenguins.
-
-```{r}
-#| label: fig-bills
-#| fig-cap: Bills of penguins
-#| echo: false
-
-ggplot(penguins, aes(x = island, fill = species)) +
-  geom_bar(alpha = 0.8) +
-  scale_fill_manual(values = c("darkorange","purple","cyan4"),
-                    guide = "none") +
-  theme_minimal() +
-  facet_wrap(~species, ncol = 1) +
-  coord_flip()
-```
+Some of our data is of penguins (), from .
 
 Talk more about it.
-
-And also planes (@fig-planes).
-(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)
-
-```{r}
-#| label: fig-planes
-#| fig-cap: Relationship between wing length and width
-#| echo: false
-#| warning: false
-#| message: false
-
-analysis_data <- read_csv(here::here("data/02-analysis_data/analysis_data.csv"))
-
-analysis_data |> 
-  ggplot(aes(x = width, y = length)) +
-  geom_point(alpha = 0.8) +
-  theme_minimal() +
-  labs(x = "Wing width (mm)",
-       y = "Wing length (mm)")
-```
 
@@
 
-Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].
+Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix ].
 
@@
 
-We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
+We run the model in R
 
@@
 
-Our results are summarized in @tbl-modelresults.
-
-```{r}
-#| echo: false
-#| eval: true
-#| warning: false
-#| message: false
-
-library(rstanarm)
-
-first_model <-
-  readRDS(file = here::here("models/first_model.rds"))
-```
-
-```{r}
-#| echo: false
-#| eval: true
-#| label: tbl-modelresults
-#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
-#| warning: false
-
-modelsummary::modelsummary(
-  list(
-    "First model" = first_model
-  ),
-  statistic = "mad",
-  fmt = 2
-)
-```
+Our results are summar
 
@@
 
-In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...
-
-In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...
-
-```{r}
-#| eval: true
-#| echo: false
-#| message: false
-#| warning: false
-#| label: fig-ppcheckandposteriorvsprior
-#| layout-ncol: 2
-#| fig-cap: "Examining how the model fits, and is affected by, the data"
-#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]
-
-pp_check(first_model) +
-  theme_classic() +
-  theme(legend.position = "bottom")
-
-posterior_vs_prior(first_model) +
-  theme_minimal() +
-  scale_color_brewer(palette = "Set1") +
-  theme(legend.position = "bottom") +
-  coord_flip()
-```
-
 ## Diagnostics
 
-@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...
+\newpage
 
-@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...
+# Appendix {#sec-appendix .unnumbered}
 
-```{r}
-#| echo: false
-#| eval: true
-#| message: false
-#| warning: false
-#| label: fig-stanareyouokay
-#| fig-cap: "Checking the convergence of the MCMC algorithm"
-#| fig-subcap: ["Trace plot", "Rhat plot"]
-#| layout-ncol: 2
+**Data Cleaning Steps**
 
-plot(first_model, "trace")
+Raw data from Open Data Toronto were prepared, so we did not need to perform many cleaning steps. A few cleaning steps are as follows:
 
-plot(first_model, "rhat")
-```
+- Recode missing values for genders other than male or female.
+- Standardize column names.
+- After realizing that there were too many missing values for other genders (over 90%), we dropped that variable.
+- Create a new Time variable to represent the first day of each recorded month in date format.
 

### 2024-12-02 paper/paper.qmd
DataAndMethodUpdate
update data and methods at second time

@@
   Column = c(
-    "_id", "EVENT_UNIQUE_ID", "REPORT_DATE", "OCC_DATE", "REPORT_YEAR",
+    "X_id", "EVENT_UNIQUE_ID", "REPORT_DATE", "OCC_DATE", "REPORT_YEAR",
     "REPORT_MONTH", "REPORT_DAY", "REPORT_DOY", "REPORT_DOW", "REPORT_HOUR",

### 2024-12-02 scripts/00-simulate_data.R

@@ -1,52 +1,41 @@
 #### Preamble ####
-# Purpose: Simulates a dataset of Australian electoral divisions, including the 
-  #state and party that won each division.
-# Author: Rohan Alexander
-# Date: 26 September 2024
-# Contact: rohan.alexander@utoronto.ca
+# Purpose: Simulates a dataset of car theft records
+# Author: Yiyue Deng
+# Date: 12/01/2024
+# Contact: yiyue.deng@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: The `tidyverse` package must be installed
-# Any other information needed? Make sure you are in the `starter_folder` rproj
+# Pre-requisites: Packages under workspace setup must be installed
+# This script simulates a dataset that looks like a subset of car theft data from open data Toronto
+# Refer to https://github.com/vanessadyy/Car_Theft for more details
 
 
 #### Workspace setup ####
 library(tidyverse)
+Sys.setlocale("LC_ALL", "en_US.UTF-8")
 set.seed(853)
 
 
-#### Simulate data ####
-# State names
-states <- c(
-  "New South Wales",
-  "Victoria",
-  "Queensland",
-  "South Australia",
-  "Western Australia",
-  "Tasmania",
-  "Northern Territory",
-  "Australian Capital Territory"
+
+#### Simulate data of size sim_size####
+sim_size <- 100
+simulated_data <- tibble(
+  OCC_DATE = as.Date("2014-01-01") + sample(0:(as.Date("2024-09-30") - as.Date("2014-01-01")), sim_size, replace = TRUE),
+  OCC_YEAR = as.integer(format(OCC_DATE, "%Y")),
+  OCC_MONTH = as.integer(format(OCC_DATE, "%m")),
+  OCC_DAY = as.integer(format(OCC_DATE, "%d")),
+  OCC_DOW = weekdays(OCC_DATE),
+  OCC_HOUR = sample(0:23, sim_size, replace = TRUE),
+  DIVISION = sample(c("D11", "D12", "D13", "D14", "D22", "D23", "D31", "D32", "D33", 
+                      "D41", "D42", "D43", "D51", "D52", "D53", "D54", "D55"), sim_size, replace = TRUE),
+  LONG_WGS84 = runif(sim_size, -79.64, -79.12),
+  LAT_WGS84 = runif(sim_size, 43.59, 43.85),
[…]

### 2024-12-02 scripts/01-test_simulated_data.R

@@ -1,20 +1,24 @@
 #### Preamble ####
-# Purpose: Tests the structure and validity of the simulated Australian 
-  #electoral divisions dataset.
-# Author: Rohan Alexander
-# Date: 26 September 2024
-# Contact: rohan.alexander@utoronto.ca
+# Purpose: Tests the structure and validity of the simulated car theft data of Toronto 
+# Author: Yiyue Deng
+# Date: 12/01/2024
+# Contact: yiyue.deng@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: 
-  # - The `tidyverse` package must be installed and loaded
-  # - 00-simulate_data.R must have been run
-# Any other information needed? Make sure you are in the `starter_folder` rproj
+# Pre-requisites: Packages under workspace setup must be installed
+# Pre-requisites: 00-simulate_data.R must have been run
+# Refer to https://github.com/vanessadyy/Car_Theft for more details
 
 
 #### Workspace setup ####
 library(tidyverse)
+library(rstudioapi)
 
-analysis_data <- read_csv("data/00-simulated_data/simulated_data.csv")
+#### read simulated data ####
+
+fd <- paste0(dirname(rstudioapi::getActiveDocumentContext()$path), "/../")  
+setwd(fd)
+
+analysis_data <- read_csv("./data/simulated_data/simulated_data.csv")
 
 # Test if the data was successfully loaded
 if (exists("analysis_data")) {
@@ -26,45 +30,60 @@ if (exists("analysis_data")) {
 
 #### Test data ####
 
-# Check if the dataset has 151 rows
-if (nrow(analysis_data) == 151) {
-  message("Test Passed: The dataset has 151 rows.")
+# Check if the dataset has 100 rows
+if (nrow(analysis_data) == 100) {
+  message("Test Passed: The dataset has 100 rows.")
 } else {
-  stop("Test Failed: The dataset does not have 151 rows.")
+  stop("Test Failed: The dataset does not have 100 rows.")
 }
 
-# Check if the dataset has 3 columns
-if (ncol(analysis_data) == 3) {
-  message("Test Passed: The dataset has 3 columns.")
+# Check if the dataset has 13 columns
+if (ncol(analysis_data) == 13) {
+  message("Test Passed: The dataset has 13 columns.")
 } else {
-  stop("Test Failed: The dataset does not have 3 columns.")
+  stop("Test Failed: The dataset does not have 13 columns.")
 }
 
-# Check if all values in the 'division' column are unique
[…]

### 2024-12-02 scripts/02-download_data.R

@@ -1,26 +1,38 @@
 #### Preamble ####
-# Purpose: Downloads and saves the data from [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Downloads and saves the data from Open Data Toronto
+# Author: Yiyue Deng
+# Date: 12/01/2024
+# Contact: yiyue.deng@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
-
+# Pre-requisites: Packages under workspace setup must be installed
+# This script is used to download data for a follow-up study of the Canadian government’s National Action Plan to Combat Car Theft
+# Refer to https://github.com/vanessadyy/Car_Theft for more details
 
 #### Workspace setup ####
 library(opendatatoronto)
 library(tidyverse)
-# [...UPDATE THIS...]
+library(rstudioapi)
 
 #### Download data ####
-# [...ADD CODE HERE TO DOWNLOAD...]
+# get package
+package <- show_package("1fc65d1e-7dae-4766-98dd-3b172e40a089")
+package
 
+# get all resources for this package
+resources <- list_package_resources("1fc65d1e-7dae-4766-98dd-3b172e40a089")
+
+# identify datastore resources; by default, Toronto Open Data sets datastore resource format to CSV for non-geospatial and GeoJSON for geospatial resources
+# here I used csv format data
+datastore_resources <- filter(resources, tolower(format) %in% c('csv'))
+
+
+# load the first datastore resource as a sample
+data <- filter(datastore_resources, row_number()==1) %>% get_resource()
+data
 
 
 #### Save data ####
-# [...UPDATE THIS...]
 # change the_raw_data to whatever name you assigned when you downloaded it.
-write_csv(the_raw_data, "inputs/data/raw_data.csv") 
-
-         
+fd <- paste0(dirname(rstudioapi::getActiveDocumentContext()$path), "/../")  
+setwd(fd)
+write_csv(data, "./data/raw_data/raw_data.csv") 

### 2024-12-02 scripts/03-clean_data.R

@@ -1,44 +1,41 @@
 #### Preamble ####
-# Purpose: Cleans the raw plane data recorded by two observers..... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 6 April 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Cleans the raw car theft data
+# Author:  Yiyue Deng
+# Date: 12/01/2024
+# Contact: yiyue.deng@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: Packages under workspace setup must be installed
+# This script is used to clean raw car theft data downloaded directly from open data Toronto
+# Refer to https://github.com/vanessadyy/Car_Theft for more details
 
 #### Workspace setup ####
 library(tidyverse)
+library(rstudioapi)
+library(janitor)
 
+# Set data directory and read raw data
+fd <- paste0(dirname(rstudioapi::getActiveDocumentContext()$path), "/../")  
+setwd(fd)
+raw_data <- read_csv("./data/raw_data/raw_data.csv")
+names(raw_data)
 #### Clean data ####
-raw_data <- read_csv("inputs/data/plane_data.csv")
-
+# Only variables of interest are retained. Please refer to the paper for the list of variables of interest and reasons.
+# Only records occurred after 2014 were retained
+# Any records with missing data were excluded
 cleaned_data <-
-  raw_data |>
-  janitor::clean_names() |>
-  select(wing_width_mm, wing_length_mm, flying_time_sec_first_timer) |>
-  filter(wing_width_mm != "caw") |>
+  raw_data %>%
+  select(X_id,	OCC_DATE,	OCC_YEAR,	OCC_MONTH,	
+         OCC_DAY,	OCC_DOW,	OCC_HOUR,	DIVISION,	
+         PREMISES_TYPE,	UCR_CODE,	OFFENCE,	
+         LONG_WGS84,	LAT_WGS84) %>%
+  filter(OCC_YEAR > 2013,
+         OCC_YEAR != "None") %>%
   mutate(
-    flying_time_sec_first_timer = if_else(flying_time_sec_first_timer == "1,35",
-                                   "1.35",
-                                   flying_time_sec_first_timer)
-  ) |>
[…]

### 2024-12-02 scripts/04-test_analysis_data.R

@@ -1,69 +1,77 @@
 #### Preamble ####
-# Purpose: Tests... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 26 September 2024 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Tests the structure and validity of the download and cleaned car theft data of Toronto 
+# Author: Yiyue Deng
+# Date: 12/01/2024
+# Contact: yiyue.deng@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: Packages under workspace setup must be installed
+# Pre-requisites: 02-download_data.R and 03-clean_data.R must have been run
+# Refer to https://github.com/vanessadyy/Car_Theft for more details
 
 
 #### Workspace setup ####
 library(tidyverse)
-library(testthat)
+library(rstudioapi)
 
-data <- read_csv("data/02-analysis_data/analysis_data.csv")
-
+#### read data ####
+fd <- paste0(dirname(rstudioapi::getActiveDocumentContext()$path), "/../")  
+setwd(fd)
+analysis_data <- read_csv("./data/analysis_data/analysis_data.csv")
 
 #### Test data ####
-# Test that the dataset has 151 rows - there are 151 divisions in Australia
-test_that("dataset has 151 rows", {
-  expect_equal(nrow(analysis_data), 151)
-})
 
-# Test that the dataset has 3 columns
-test_that("dataset has 3 columns", {
-  expect_equal(ncol(analysis_data), 3)
-})
+# Check if the dataset has 13 columns
+if (ncol(analysis_data) == 13) {
+  message("Test Passed: The dataset has 13 columns.")
+} else {
+  stop("Test Failed: The dataset does not have 13 columns.")
+}
 
-# Test that the 'division' column is character type
-test_that("'division' is character", {
-  expect_type(analysis_data$division, "character")
-})
+# Check if all values in the 'X_id' column are unique
+if (n_distinct(analysis_data$X_id) == nrow(analysis_data)) {
+  message("Test Passed: All values in 'X_id' are unique.")
+} else {
+  stop("Test Failed: The 'X_id' column contains duplicate values.")
+}
 
-# Test that the 'party' column is character type
-test_that("'party' is character", {
-  expect_type(analysis_data$party, "character")
-})
[…]

### 2024-12-02 scripts/05-exploratory_data_analysis.R

@@ -1,19 +1,22 @@
 #### Preamble ####
-# Purpose: Models... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Exploratory analysis of car theft data from open data Toronto
+# Author: Yiyue Deng
+# Date: 12/01/2024
+# Contact: yiyue.deng@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: Packages under workspace setup must be installed
+# Pre-requisites: 02-download_data.R and 03-clean_data.R must have been run
+# Refer to https://github.com/vanessadyy/Car_Theft for more details
 
 
 #### Workspace setup ####
 library(tidyverse)
-library(rstanarm)
+library(rstudioapi)
 
 #### Read data ####
-analysis_data <- read_csv("data/analysis_data/analysis_data.csv")
+fd <- paste0(dirname(rstudioapi::getActiveDocumentContext()$path), "/../")  
+setwd(fd)
+analysis_data <- read_csv("./data/analysis_data/analysis_data.csv")
 
 ### Model data ####
 first_model <-

### 2024-12-04 paper/paper.qmd
wholeupdate
wholeupdate

--- 2024-12-01 paper/paper.qmd
@@
 date-format: long
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
+abstract: "Vehicle theft has become a significant public concern in Canada, highlighting both systemic challenges and promising governmental interventions.
+This study evaluates the effectiveness of the 2024 National Action Plan, which led to a 16.9% reduction in the mean monthly count of car thefts in Toronto.
+Spatial and temporal analyses reveal variations across police divisions and premise types, emphasizing the need for targeted strategies.
+Despite encouraging trends, organized theft networks and data limitations underscore the need for further research and adaptive policymaking."
 include-before:
@@
 library(tidyverse)
-library(tidyverse)
 library(here)
@@
 library(kableExtra)
+library(distributions3)
+library(topmodels)
+library(MASS)
+library(sf)
+library(rstudioapi)
+library(opendatatoronto)
+library(janitor)
 ```
@@
 
+## Estimand
+
+The primary estimand of this study is the monthly reduction in mean counts of car theft cases in Toronto following the implementation of the National Action Plan.
+This is defined as the difference in the mean monthly car theft count before and after January 2024, when the plan starts.
+The estimand will be evaluated across all vehicle theft records reported in Toronto, stratified by key covariates.
+This analysis assumes that the observed trends are not confounded by unrelated temporal factors.
+
 # Data {#sec-data}
@@
     "Location Type of Offence",
-    "Premises Type of Offence",
+    "premise types of Offence",
     "UCR Code for Offence",
@@
 
-We used R 4.3 for analysis and followed a public template @citeR@citeRohan. and a collection of R packages were used in this study:
+We used R 4.3 for analysis and followed a public template @citeR@citeRohan, and a collection of R packages were used in this study:
 
@@
 - knitr @Citeknitr
-- kableExtra
+- kableExtra @citekableExtra
+- topmodels @citetopmodels
+- distributions3 @citedistributions3
+- MASS @citemass
+- sf @citesf
+- rstudioapi @citerstudioapi
+- opendatatoronto @citeopendatatoronto
+- janitor @citejanitor
 
@@
 Data cleaning steps are described in the [Appendix](#sec-appendix).
-A sample of the cleaned data is presented in Table 2-4.
+A sample set of the cleaned data is presented in Table 2-4.
 Among all variables, Year, Total_Death, Male, and Female are coded as integers, Month is coded as abbreviations in character format, and Time is coded in date format.
@@
 
-Estimand paragraph
+# Methods {#sec-methods}
 
-Results paragraph
+## Poisson Regression
 
-Why it matters paragraph
+The Poisson distribution is utilized to model the monthly count of car thefts, with a binary variable "action" serving as an indicator of whether the time of theft occurrence falls after the implementation of the national action plan.
+The following are the formulas for the Poisson regression model:
 
-Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....
+$$
+y_i \sim \text{Poisson}(\lambda_i)
+$$
 
-# Data {#sec-data}
+$$
+\log(\lambda_i) = \beta_0 + \beta_1 \cdot \text{time} + \beta_2 \cdot \text{action} + \cdots + \beta_k \cdot \text{covariate}_k
+$$
+where y_i is the count of car theft occurance.
 
-## Overview
+We ran a generalized linear model with a log link function, ensuring that all of the predicted values will be positive, and using a Poisson error distribution.
 
-We use the statistical programming language R .... Our data .... Following , we consider...
+## Quasi-likelihood Poisson Regression
 
-Overview text
+To address the potential for overdispersion, we also performed a quasi-likelihood Poisson regression.
+This approach uses the same mean function as standard Poisson regression but estimates parameters using quasi-maximum likelihood estimation or generalized estimating equations, with adjustments for an estimated dispersion parameter.
+We compared the results from Poisson and Quasi-Poisson regression models, selecting the more suitable model based on a comprehensive diagnostic evaluation.
+This assessment included checks for overdispersion, identification of outliers, and analysis of residual distribution.
+Further details are provided in the [results section]{@sec-results}.
 
-## Measurement
+# Results {#sec-results}
 
-Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.
+## Descriptives
 
-## Outcome variables
+```{r}
+#| label: fig-quarterly-car-theft
+#| echo: false
+#| warning: false
+#| fig-cap: "Histogram of quarterly count of car theft in Toronto"
 
-Add graphs, tables and text. Use sub-sub-headings for each outcome variable or update the subheading to be singular.
+quarterly_counts <- dt %>%
+  mutate(
+    YEARQUARTER = paste0(format(OCC_DATE, "%Y"), "-Q", ceiling(as.numeric(format(OCC_DATE, "%m")) / 3))  # Extract year and quarter
+  ) %>%
+  group_by(YEARQUARTER) %>%
+  summarise(Crime_Count = n(), .groups = "drop") %>%
+  mutate(
+    time = as.numeric(as.Date(paste0(substr(YEARQUARTER, 1, 4), "-", (as.numeric(substr(YEARQUARTER, 7, 7)) - 1) * 3 + 1, "-01"))),
+    action = ifelse(time >= as.numeric(as.Date("2024-01-01")), 1, 0)
+  )
 
-Some of our data is of penguins (), from .
+ggplot(quarterly_counts, aes(x = YEARQUARTER, y = Crime_Count)) +
+  geom_col(fill = "steelblue", color = "black") +
+  labs(
+    x = "Month",
+    y = "Number of Cases"
+  ) +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+```
 
-Talk more about it.
+Before running the model, we firstly draw a few figures and tables as descriptives to have a quickview of the data.
 
-Talk way more about it.
+### Timely Trend (Quarterly)
 
-## Predictor variables
+@fig-quarterly-car-theft illustrates the overall trend in the quarterly count of car theft cases in Toronto since 2014.
+The number of cases generally fluctuated around 2,000 per quarter.
+Notable periods of lower counts include 2016 and the first two quarters of 2021, while higher counts were observed during 2020 and the period from 2022 to 2023.
+A decline is also apparent after 2024, though further evidence is needed to confirm this trend.
 
-Add graphs, tables and text.
+### Spatial View
 
-Use sub-sub-headings for each outcome variable and feel free to combine a few into one if they go together naturally.
+```{r}
+#| label: fig-Spatial-View
+#| echo: false
+#| warning: false
+#| fig-cap: "Spatial view of vehicle theft cases by division for September 2024"
 
-# Model
+division_sf <- st_as_sf(dt %>% filter(OCC_YEAR==2024, OCC_MONTH=="September"), coords = c("LONG_WGS84", "LAT_WGS84"), crs = 4326) %>%
+  group_by(DIVISION) %>%
+  summarise(geometry = st_combine(geometry)) %>%
+  st_cast("POLYGON")
+ggplot(data = division_sf) +
+  geom_sf(aes(fill = factor(DIVISION)), color = "black") +
+  scale_fill_viridis_d(name = "Division ID") +
+  labs( x = "Longitude", y = "Latitude") +
+  theme_minimal()
+```
 
-The goal of our modelling strategy is twofold. Firstly,...
+@fig-Spatial-View illustrates the spatial distribution of car theft occurrences in September 2024 across police divisions.
+@fig-division-chart displays the overall count of car theft cases by division.
+Notably, Division D54 reported the lowest number of car theft cases, with Divisions D11 to D13 also showing relatively low counts, while Division D33 recorded the highest.
+Further analysis is needed, considering additional information about public safety conditions within these division areas.
 
-Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix ].
+```{r}
+#| label: fig-division-chart
+#| echo: false
+#| warning: false
+#| fig-cap: "Barchart of count of occurrences by divisions"
 
-## Model set-up
+ggplot(dt, aes(x = DIVISION)) +
+  geom_bar(fill = "steelblue", color = "black") +
+  labs(
+    x = "DIVISION",
+    y = "Number of Occurrences"
+  ) +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+```
 
-Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.
+### Premise Types
 
-\begin{align} 
-y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
-\mu_i &= \alpha + \beta_i + \gamma_i\\
-\alpha &\sim \mbox{Normal}(0, 2.5) \\
-\beta &\sim \mbox{Normal}(0, 2.5) \\
-\gamma &\sim \mbox{Normal}(0, 2.5) \\
-\sigma &\sim \mbox{Exponential}(1)
-\end{align}
+@fig-premises-chart reveals that car theft cases occurred most frequently outside, followed by educational institutions, while transit locations reported the fewest cases.
+Across all premise types, two periods of increase were observed in 2019 and 2022-2023, with a decline noted after 2024.
 
-We run the model in R
+```{r}
+#| label: fig-premises-chart
+#| echo: false
+#| warning: false
+#| fig-cap: "Linechart of count of occurrences by premise types and year"
 
-### Model justification
+table_data <- table(dt$PREMISES_TYPE, dt$OCC_YEAR)
+df_table <- as.data.frame(table_data)
+colnames(df_table) <- c("PREMISES_TYPE", "OCC_YEAR", "Count")
+df_table$OCC_YEAR <- as.numeric(as.character(df_table$OCC_YEAR))
+ggplot(df_table, aes(x = OCC_YEAR, y = Count, color = PREMISES_TYPE, group = PREMISES_TYPE)) +
+  geom_line(size = 1) +
+  geom_point(size = 2) +
+  labs(
+    x = "Year",
+    y = "Count of Cases",
+    color = "premise types"
+  ) +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+```
 
-We expect a positive relationship between the size of the wings and time spent aloft. In particular...
+### Offence Types
 
-We can use maths by including latex between dollar signs, for instance $\theta$.
+@fig-offence-chart shows that most car theft cases were categorized as "Theft From Motor Vehicle Under," involving financial losses of less than $5,000.
+The number of severe cases remained relatively stable over the years, while lighter cases experienced two increases in 2019 and 2022, with declines observed in 2016 and after 2024.
 
-# Results
+```{r}
+#| label: fig-offence-chart
+#| echo: false
+#| warning: false
+#| fig-cap: "Linechart of count of occurrences by offence types and year"
 
-Our results are summar
+offence_table <- table(dt$OFFENCE, dt$OCC_YEAR)
+df_offence <- as.data.frame(offence_table)
+colnames(df_offence) <- c("OFFENCE", "OCC_YEAR", "Count")
+df_offence$OCC_YEAR <- as.numeric(as.character(df_offence$OCC_YEAR))
+ggplot(df_offence, aes(x = OCC_YEAR, y = Count, color = OFFENCE, group = OFFENCE)) +
+  geom_line(size = 1) +
+  geom_point(size = 2) +
+  labs(
+    x = "Year",
+    y = "Count of Cases",
+    color = "Offence Type"
+  ) +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+```
+
+### Day of the Week
+
+@fig-dow-chart indicates that car theft crimes were distributed relatively evenly across all days of the week, with a slight increase observed on Fridays.
+
+```{r}
+#| label: fig-dow-chart
+#| echo: false
+#| warning: false
+#| fig-cap: "Barchart of count of occurrences by day of the week"
+
+ggplot(dt, aes(x = OCC_DOW)) +
+  geom_bar(fill = "steelblue", color = "black") +
+  labs(
+    x = "Day of the Week",
+    y = "Number of Occurrences"
+  ) +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+```
+
+### Day of the Month
+
+Interestingly, @fig-day-chart reveals that car theft cases occurred most frequently on the first day of the month and least often at the end of the month.
+The difference could be as high as 1,000 cases over a 10-year period, equivalent to approximately 30%.
+
+```{r}
+#| label: fig-day-chart
+#| echo: false
+#| warning: false
+#| fig-cap: "Barchart of count of occurrences by day of the month"
+
+ggplot(dt, aes(x = OCC_DAY)) +
+  geom_bar(fill = "steelblue", color = "black") +
+  labs(
+    x = "Day of the Month",
+    y = "Number of Occurrences"
+  ) +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+```
+
+## Poisson Model
+
+### Point Estimates
+
+@tbl-Poisson indicates that the variable "action" had a significant effect on the mean count of car theft cases.
+Specifically, after the implementation of the national action plan, there was an approximate 1−exp(−0.185)=1−0.831=16.9% reduction in the mean count of car theft cases.
+This result highlights the potential effectiveness of the action plan in mitigating car theft, suggesting a notable decrease in incidents following its introduction.
+
+```{r}
+#| label: tbl-Poisson
+#| echo: false
+#| warning: false
+#| tbl-cap: "Poisson regression results"
+
+monthly_counts_all <- dt %>%
+  mutate(
+    YEARMONTH = format(OCC_DATE, "%Y-%m")  # Extract year and month in "YYYY-MM" format
+  ) %>%
+  group_by(YEARMONTH) %>%
+  summarise(Crime_Count = n(), .groups = "drop") %>%
+  mutate(
+    time = as.numeric(as.Date(paste0(YEARMONTH, "-01"))),
+    action = ifelse(time >= as.numeric(as.Date("2024-01-01")),1,0))
+poisson_model_all <- glm(
+  Crime_Count ~ time +action,
+  family = poisson(link = "log"),
+  data = monthly_counts_all
+)
+kable(summary(poisson_model_all)$coefficients)
+```
+
+### Prediction
+
+Predictions from the model in @fig-Poisson-pred demonstrate a significant difference in the expected monthly car theft counts, with post-2024 counts being approximately 100 cases lower per month on average.
+This reduction suggests a measurable impact of national action plan starting in 2024.
+
+```{r}
+#| label: fig-Poisson-pred
+#| echo: false
+#| warning: false
+#| fig-cap: "Poisson regression prediction"
+
+monthly_counts_all <- monthly_counts_all %>%
+  mutate(
+    Predicted_Count = predict(poisson_model_all, type = "response")
+  )
+ggplot(monthly_counts_all, aes(x = as.Date(time))) +
+  geom_line(aes(y = Crime_Count, color = "Actual", linetype = "Actual"), size = 1) + # Actual counts
+  geom_line(aes(y = Predicted_Count, color = "Predicted", linetype = "Predicted"), size = 1) + # Predicted counts
+  scale_color_manual(values = c("Actual" = "blue", "Predicted" = "red")) + # Define line colors
+  scale_linetype_manual(values = c("Actual" = "solid", "Predicted" = "dashed")) + # Define line types
+  labs(
+    x = "Time",
+    y = "Car Theft Crime Count",
+    color = "Type", # Legend title for color
+    linetype = "Type" # Legend title for linetype
+  ) +
+  theme_minimal() +
+  theme(legend.position = "right")
+```
+
+### Model Diagnosis
+
+The dispersion factor from @tbl-Poisson-dispersion is 14.74, indicating significant overdispersion in the data.
+The rootogram in @fig-Poisson-rootogram reveals overfitting for crime counts in the range of 700–800 and underfitting for other ranges.
+Diagnostic plots in @fig-Poisson-diagnosis show no major deflation in the QQ plot, suggesting a normal distribution of residuals.
+Leverage plots indicate only three significant outliers, corresponding to observations 5, 26, and 122.
+
+```{r}
+#| label: tbl-Poisson-dispersion
+#| echo: false
+#| warning: false
+#| tbl-cap: "Poisson regression dispersion factor"
+
+dispersion_stat <- sum(residuals(poisson_model_all, type = "pearson")^2) / poisson_model_all$df.residual
+kable(data.frame(Stats="Dispersion Factor",Value=dispersion_stat))
+```
+
+```{r}
+#| label: fig-Poisson-rootogram
+#| echo: false
+#| warning: false
+#| fig-cap: "Poisson regression rootogram"
+
+rootogram(poisson_model_all)
+```
+
+```{r}
+#| label: fig-Poisson-diagnosis
+#| echo: false
+#| warning: false
+#| fig-cap: "Poisson regression model diagnosis"
+
+par(mfrow = c(2, 2)) 
+plot(poisson_model_all)
+```
+
+## Quasi-Poisson Model
+
+@tbl-Q-Poisson shows that the Quasi-Poisson model produced the same point estimate for the effect of "action," indicating a 16.9% reduction in the mean count of car theft cases following the implementation of the national action plan.
+However, the Quasi-Poisson model estimated a larger standard error, resulting in narrower confidence intervals (see @tbl-Q-Poisson), which may provide a more reliable and appropriate measure of precision.
+
+```{r}
+#| label: tbl-Q-Poisson
+#| echo: false
+#| warning: false
+#| tbl-cap: "Quasi-likelihood Poisson regression results"
+
+poisson_model_all2 <- glm(Crime_Count ~ time +action,
+                          family = quasipoisson(),
+                          data = monthly_counts_all)
+
+kable(summary(poisson_model_all2)$coefficients)
+```
+
+```{r}
+#| label: tbl-Q-P-CI
+#| echo: false
+#| warning: false
+#| tbl-cap: "Comparison between Quasi-likelihood Poisson and Poisson estimated confidence intervals"
+
+t1 <- as.data.frame(confint(poisson_model_all))
+t2 <- as.data.frame(confint(poisson_model_all2))
+t1$model_type <- "Poisson Model"
+t2$model_type <- "Quasi-Poisson Model"
+kable(rbind(t1,t2))
+```
+
+## Model with covariates
+
+```{r}
+#| label: tbl-Q-P-PREMISES_TYPE
+#| echo: false
+#| warning: false
+#| tbl-cap: "Quasi-likelihood Poisson regression with covariate of premise types"
+monthly_counts_PREMISES_TYPE <- dt %>%
+  mutate(
+    YEARMONTH = format(OCC_DATE, "%Y-%m")  # Extract year and month in "YYYY-MM" format
+  ) %>%
+  group_by(YEARMONTH,PREMISES_TYPE) %>%
+  summarise(Crime_Count = n(), .groups = "drop") %>%
+  mutate(
+    time = as.Date(paste0(YEARMONTH, "-01")),
+    action = ifelse(time >= as.Date("2024-01-01"),1,0))
+poisson_model_PREMISES_TYPE <- glm(
+  Crime_Count ~ time+PREMISES_TYPE +action,
+  family = quasipoisson(),
+  data = monthly_counts_PREMISES_TYPE
+)
+kable(summary(poisson_model_PREMISES_TYPE)$coefficients)
+```
+
+```{r}
+#| label: tbl-Q-P-OFFENCE
+#| echo: false
+#| warning: false
+#| tbl-cap: "Quasi-likelihood Poisson regression with covariate of offence types"
+monthly_counts_OFFENCE <- dt %>%
+  mutate(
+    YEARMONTH = format(OCC_DATE, "%Y-%m")  # Extract year and month in "YYYY-MM" format
+  ) %>%
+  group_by(YEARMONTH,OFFENCE) %>%
+  summarise(Crime_Count = n(), .groups = "drop") %>%
+  mutate(
+    time = as.Date(paste0(YEARMONTH, "-01")),
+    action = ifelse(time >= as.Date("2024-01-01"),1,0))
+poisson_model_OFFENCE <- glm(
+  Crime_Count ~ time+OFFENCE +action,
+  family = quasipoisson(),
+  data = monthly_counts_OFFENCE
+)
+kable(summary(poisson_model_OFFENCE)$coefficients)
+```
+
+```{r}
+#| label: tbl-Q-P-DIVISION
+#| echo: false
+#| warning: false
+#| tbl-cap: "Quasi-likelihood Poisson regression with covariate of divisions"
+monthly_counts_DIVISION <- dt %>%
+  mutate(
+    YEARMONTH = format(OCC_DATE, "%Y-%m")  # Extract year and month in "YYYY-MM" format
+  ) %>%
+  group_by(YEARMONTH,DIVISION) %>%
+  summarise(Crime_Count = n(), .groups = "drop") %>%
+  mutate(
+    time = as.Date(paste0(YEARMONTH, "-01"))  ,
+    action = ifelse(time >= as.Date("2024-01-01"),1,0))
+poisson_model_DIVISION <- glm(
+  Crime_Count ~ time+DIVISION +action,
+  family = quasipoisson(),
+  data = monthly_counts_DIVISION
+)
+kable(summary(poisson_model_DIVISION)$coefficients)
+```
+
+@tbl-Q-P-PREMISES_TYPE, @tbl-Q-P-OFFENCE, and @tbl-Q-P-DIVISION present point estimates from models incorporating different covariates.
+The results indicate that parking outside significantly increases the risk of car theft occurrences.
+Additionally, theft categorized as "Theft From Motor Vehicle Under" was associated with a higher number of car theft cases.
+Among the divisions, Division 32 exhibited the highest risk of car theft occurrence.
 
@@
 
-## First discussion point {#sec-first-point}
+## Impact of the National Action Plan {#sec-first-point}
 
-If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
+The results of this study indicate that the National Action Plan implemented in early 2024 has shown a measurable impact on reducing car theft in Toronto.
+Our analysis demonstrates a 16.9% reduction in the mean count of car theft cases after the action plan’s introduction, as evidenced by both Poisson and Quasi-Poisson regression models.
+This decline is further supported by the observed decrease in monthly crime counts after 2024, with an average reduction of 100 cases per month.
+These findings highlight the potential effectiveness of government-led initiatives in curbing auto theft, although additional time is needed to assess the plan’s long-term impact comprehensively.
 
-## Second discussion point
+## Spatial and Temporal Patterns of Car Theft {#sec-second-point}
 
-Please don't use these as sub-heading labels - change them to be what your point actually is.
+The spatial analysis revealed substantial variation in car theft occurrences across police divisions.
+Division 32 consistently exhibited the highest risk of car theft, while Divisions 11, 12, 13, and 54 reported the lowest counts.
+These patterns may reflect underlying differences in socioeconomic conditions, policing strategies, or community vigilance within these divisions.
+Additionally, the analysis of premise types indicated that thefts occurred most frequently in outdoor locations, suggesting that vehicle owners parking outside face higher risks.
+Temporal trends revealed spikes in crime during certain years, notably 2019 and 2022–2023, followed by a decline in 2024.
+Interestingly, car thefts were most common at the beginning of the month, possibly linked to behavioral or logistical patterns among offenders.
 
-## Third discussion point
+## Effectiveness of the National Action Plan {#sec-third-point}
 
-## Weaknesses and next steps
+While the National Action Plan appears to have contributed to a decline in car theft cases, its effectiveness likely varies by division and offense type.
+The reduction in "Theft From Motor Vehicle Under" cases aligns with the broader trend but raises questions about whether certain measures disproportionately target less severe crimes.
+Furthermore, the use of enhanced technologies and intelligence sharing could be more impactful in divisions with historically higher crime rates, such as Division 32.
+These findings underscore the importance of tailoring interventions to the specific needs and challenges of different regions.
 
-Weaknesses and next steps should also be included.
+## Weaknesses and Limitations
+
+This study has several limitations.
+First, the data used were limited to reported cases, potentially underestimating the true extent of car theft due to unreported incidents.
+Second, the overdispersion identified in the Poisson model suggests variability that may not have been fully captured by the predictors included in our analysis.
+Although the Quasi-Poisson model provided more appropriate standard error estimates, additional covariates, such as socioeconomic factors or policing intensity, could further improve model accuracy.
+Lastly, the relatively short timeframe since the National Action Plan’s implementation restricts our ability to evaluate its sustained effects.
+
+## Recommendations and Future Work
+
+Based on the findings, we recommend that policymakers focus on areas with persistently high car theft rates, such as Division 32, to deploy targeted interventions.
+Efforts to improve public awareness of high-risk premise types, such as outdoor parking, could also reduce theft occurrences.
+Further research should explore the role of socioeconomic and environmental factors in shaping car theft trends, leveraging longer-term data to evaluate the sustained effectiveness of the National Action Plan.
+Additionally, integrating machine learning approaches could enhance predictive modeling and offer more granular insights into crime dynamics.
 
@@
 
-# Additional data details
-
-# Model details {#sec-model-details}
-
-## Posterior predictive check
-
-## Diagnostics
-
-\newpage
-
-# Appendix {#sec-appendix .unnumbered}
-
 **Data Cleaning Steps**
@@
 
-- Recode missing values for genders other than male or female.
-- Standardize column names.
-- After realizing that there were too many missing values for other genders (over 90%), we dropped that variable.
-- Create a new Time variable to represent the first day of each recorded month in date format.
+- Only variables of interest were kept.
+- All records with all types of missing were excluded.
+- Only record indicating car theft cases occurrence after 2014 were kept.
 

### 2024-12-04 scripts/05-exploratory_data_analysis.R

@@ -1,5 +1,5 @@
 #### Preamble ####
-# Purpose: Exploratory analysis of car theft data from open data Toronto
+# Purpose: First try of modeling car theft data from open data Toronto
 # Author: Yiyue Deng
 # Date: 12/01/2024
 # Contact: yiyue.deng@mail.utoronto.ca
@@ -10,31 +10,80 @@
 
 
 #### Workspace setup ####
-library(tidyverse)
 library(rstudioapi)
+library(MASS)
+library(tidyverse)
+library(topmodels)
+library(knitr)
+
 
 #### Read data ####
 fd <- paste0(dirname(rstudioapi::getActiveDocumentContext()$path), "/../")  
 setwd(fd)
 analysis_data <- read_csv("./data/analysis_data/analysis_data.csv")
 
-### Model data ####
-first_model <-
-  stan_glm(
-    formula = flying_time ~ length + width,
-    data = analysis_data,
-    family = gaussian(),
-    prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_aux = exponential(rate = 1, autoscale = TRUE),
-    seed = 853
+#### data transformations####
+monthly_counts_all <- analysis_data %>%
+  mutate(
+    YEARMONTH = format(OCC_DATE, "%Y-%m")  # Extract year and month in "YYYY-MM" format
+  ) %>%
+  group_by(YEARMONTH) %>%
+  summarise(Crime_Count = n(), .groups = "drop") %>%
+  mutate(
+    time = as.numeric(as.Date(paste0(YEARMONTH, "-01"))),
+    action = ifelse(time >= as.numeric(as.Date("2024-01-01")),1,0))
+write_csv(monthly_counts_all,"./data/analysis_data/monthly_counts_all.csv")
+
+####simple poisson regression####
+poisson_model_all <- glm(
+  Crime_Count ~ time +action,
+  family = poisson(link = "log"),
+  data = monthly_counts_all
+)
+
+kable(summary(poisson_model_all)$coefficients)
+kable(confint(poisson_model_all))
+
+
+#####prediction####
+
+monthly_counts_all <- monthly_counts_all %>%
+  mutate(
+    Predicted_Count = predict(poisson_model_all, type = "response")
   )
 
+##### Plot actual vs. predicted counts####
+
+ggplot(monthly_counts_all, aes(x = as.Date(time))) +
[…]

### 2024-12-04 scripts/06-model_data.R

@@ -1,37 +1,77 @@
 #### Preamble ####
-# Purpose: Models... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Second modeling process of paper analyzing car theft data from open data Toronto
+# Author: Yiyue Deng
+# Date: 12/01/2024
+# Contact: yiyue.deng@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: Packages under workspace setup must be installed
+# Pre-requisites: 02-download_data.R, 03-clean_data.R and 05-exploratory_data_analysis.R must have been run
+# Refer to https://github.com/vanessadyy/Car_Theft for more details
 
 
 #### Workspace setup ####
+library(rstudioapi)
+library(MASS)
 library(tidyverse)
-library(rstanarm)
+library(topmodels)
+library(knitr)
+
 
 #### Read data ####
-analysis_data <- read_csv("data/analysis_data/analysis_data.csv")
+fd <- paste0(dirname(rstudioapi::getActiveDocumentContext()$path), "/../")  
+setwd(fd)
+analysis_data <- read_csv("./data/analysis_data/analysis_data.csv")
+monthly_counts_all <- read_csv("./data/analysis_data/monthly_counts_all.csv")
 
-### Model data ####
-first_model <-
-  stan_glm(
-    formula = flying_time ~ length + width,
-    data = analysis_data,
-    family = gaussian(),
-    prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_aux = exponential(rate = 1, autoscale = TRUE),
-    seed = 853
+
+#read previous model#
+poisson_model_all <- readRDS(  file = "./models/first_model.rds"
+)
+
+####quasi-likelihood poisson regression####
+poisson_model_all2 <- glm(Crime_Count ~ time +action,
+                          family = quasipoisson(),
+                          data = monthly_counts_all)
+
+kable(summary(poisson_model_all2)$coefficients)
+kable(confint(poisson_model_all2))
+
+
+
+###prediction###
[…]

### 2024-12-04 scripts/07-covariate_model.R

@@ -1,0 +1,96 @@
+#### Preamble ####
+# Purpose: Add covariates of premises type, offence and division, separately, to second model of 
+# paper analyzing car theft data from open data Toronto
+# Author: Yiyue Deng
+# Date: 12/01/2024
+# Contact: yiyue.deng@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: Packages under workspace setup must be installed
+# Pre-requisites: 02-download_data.R, 03-clean_data.R must have been run
+# Refer to https://github.com/vanessadyy/Car_Theft for more details
+
+
+#### Workspace setup ####
+library(rstudioapi)
+library(MASS)
+library(tidyverse)
+library(topmodels)
+library(knitr)
+
+#### Read data ####
+fd <- paste0(dirname(rstudioapi::getActiveDocumentContext()$path), "/../")  
+setwd(fd)
+analysis_data <- read_csv("./data/analysis_data/analysis_data.csv")
+
+
+####quasi-likelihood poisson regression with premises type####
+monthly_counts_PREMISES_TYPE <- analysis_data %>%
+  mutate(
+    YEARMONTH = format(OCC_DATE, "%Y-%m")  # Extract year and month in "YYYY-MM" format
+  ) %>%
+  group_by(YEARMONTH,PREMISES_TYPE) %>%
+  summarise(Crime_Count = n(), .groups = "drop") %>%
+  mutate(
+    time = as.numeric(as.Date(paste0(YEARMONTH, "-01"))),
+    action = ifelse(time >= as.numeric(as.Date("2024-01-01")),1,0))
+poisson_model_PREMISES_TYPE <- glm(
+  Crime_Count ~ time+PREMISES_TYPE +action,
+  family = quasipoisson(),
+  data = monthly_counts_PREMISES_TYPE
+)
+kable(summary(poisson_model_PREMISES_TYPE)$coefficients)
+kable(confint(poisson_model_PREMISES_TYPE))
+
+####quasi-likelihood poisson regression with offence type####
+monthly_counts_OFFENCE <- analysis_data %>%
+  mutate(
+    YEARMONTH = format(OCC_DATE, "%Y-%m")  # Extract year and month in "YYYY-MM" format
+  ) %>%
+  group_by(YEARMONTH,OFFENCE) %>%
+  summarise(Crime_Count = n(), .groups = "drop") %>%
+  mutate(
[…]

### 2024-12-04 scripts/08-replications.R

@@ -1,0 +1,112 @@
+#### Preamble ####
+# Purpose: Replicated tables and figures of descriptive analysis
+# Author: Yiyue Deng
+# Date: 12/01/2024
+# Contact: yiyue.deng@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: Packages under workspace setup must be installed
+# Pre-requisites: 02-download_data.R, 03-clean_data.R must have been run
+# Refer to https://github.com/vanessadyy/Car_Theft for more details
+
+
+#### Workspace setup ####
+library(rstudioapi)
+library(MASS)
+library(tidyverse)
+library(topmodels)
+library(knitr)
+library(sf)
+
+#### Read data ####
+fd <- paste0(dirname(rstudioapi::getActiveDocumentContext()$path), "/../")  
+setwd(fd)
+analysis_data <- read_csv("./data/analysis_data/analysis_data.csv")
+
+### quarterly trend histogram###
+quarterly_counts <- analysis_data %>%
+  mutate(
+    YEARQUARTER = paste0(format(OCC_DATE, "%Y"), "-Q", ceiling(as.numeric(format(OCC_DATE, "%m")) / 3))  # Extract year and quarter
+  ) %>%
+  group_by(YEARQUARTER) %>%
+  summarise(Crime_Count = n(), .groups = "drop") %>%
+  mutate(
+    time = as.numeric(as.Date(paste0(substr(YEARQUARTER, 1, 4), "-", (as.numeric(substr(YEARQUARTER, 7, 7)) - 1) * 3 + 1, "-01"))),
+    action = ifelse(time >= as.numeric(as.Date("2024-01-01")), 1, 0)
+  )
+
+ggplot(quarterly_counts, aes(x = YEARQUARTER, y = Crime_Count)) +
+  geom_col(fill = "steelblue", color = "black") +
+  labs(
+    title = "Histogram of quarterly Count of Car Theft Cases",
+    x = "Month",
+    y = "Number of Cases"
+  ) +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+
+### sf plot ###
+division_sf <- st_as_sf(analysis_data %>% filter(OCC_YEAR==2024, OCC_MONTH=="September"), coords = c("LONG_WGS84", "LAT_WGS84"), crs = 4326) %>%
+  group_by(DIVISION) %>%
+  summarise(geometry = st_combine(geometry)) %>%
[…]

### 2024-12-04 scripts/00-simulate_data.R
wholepush
push before submission

@@ -8,34 +8,60 @@
 # This script simulates a dataset that looks like a subset of car theft data from open data Toronto
 # Refer to https://github.com/vanessadyy/Car_Theft for more details
 
-
 #### Workspace setup ####
-library(tidyverse)
-Sys.setlocale("LC_ALL", "en_US.UTF-8")
-set.seed(853)
+# Load the necessary library
+library(tidyverse) # For data manipulation and visualization
+Sys.setlocale("LC_ALL", "en_US.UTF-8") # Set locale to ensure compatibility with date and text formats
+set.seed(853) # Set random seed for reproducibility
 
-
-
-#### Simulate data of size sim_size####
+#### Simulate data of size sim_size ####
+# Define the size of the simulated dataset
 sim_size <- 100
+
+# Create a simulated dataset with specified fields
 simulated_data <- tibble(
+  # Generate random dates between 2014-01-01 and 2024-09-30
   OCC_DATE = as.Date("2014-01-01") + sample(0:(as.Date("2024-09-30") - as.Date("2014-01-01")), sim_size, replace = TRUE),
+  
+  # Extract year, month, and day from OCC_DATE
   OCC_YEAR = as.integer(format(OCC_DATE, "%Y")),
   OCC_MONTH = as.integer(format(OCC_DATE, "%m")),
   OCC_DAY = as.integer(format(OCC_DATE, "%d")),
+  
+  # Get the day of the week (e.g., Monday, Tuesday)
   OCC_DOW = weekdays(OCC_DATE),
+  
+  # Generate random hour of the day (0-23)
   OCC_HOUR = sample(0:23, sim_size, replace = TRUE),
+  
+  # Randomly assign one of the predefined police divisions
   DIVISION = sample(c("D11", "D12", "D13", "D14", "D22", "D23", "D31", "D32", "D33", 
                       "D41", "D42", "D43", "D51", "D52", "D53", "D54", "D55"), sim_size, replace = TRUE),
+  
+  # Generate random longitude within Toronto's geographic bounds
   LONG_WGS84 = runif(sim_size, -79.64, -79.12),
+  
+  # Generate random latitude within Toronto's geographic bounds
   LAT_WGS84 = runif(sim_size, 43.59, 43.85),
+  
+  # Randomly assign one of the premise types
[…]

### 2024-12-04 scripts/01-test_simulated_data.R

@@ -10,50 +10,51 @@
 
 
 #### Workspace setup ####
-library(tidyverse)
-library(rstudioapi)
+# Load the necessary libraries
+library(tidyverse)     # For data manipulation and visualization
+library(rstudioapi)    # For retrieving the file path dynamically
 
-#### read simulated data ####
-
+#### Read simulated data ####
+# Dynamically set the working directory to the parent folder of the current script
 fd <- paste0(dirname(rstudioapi::getActiveDocumentContext()$path), "/../")  
 setwd(fd)
 
+# Load the simulated data from the specified directory
 analysis_data <- read_csv("./data/simulated_data/simulated_data.csv")
 
 # Test if the data was successfully loaded
 if (exists("analysis_data")) {
-  message("Test Passed: The dataset was successfully loaded.")
+  message("Test Passed: The dataset was successfully loaded.") # Confirm successful loading
 } else {
-  stop("Test Failed: The dataset could not be loaded.")
+  stop("Test Failed: The dataset could not be loaded.") # Error if data not loaded
 }
-
 
 #### Test data ####
 
 # Check if the dataset has 100 rows
 if (nrow(analysis_data) == 100) {
-  message("Test Passed: The dataset has 100 rows.")
+  message("Test Passed: The dataset has 100 rows.") # Confirm row count matches
 } else {
-  stop("Test Failed: The dataset does not have 100 rows.")
+  stop("Test Failed: The dataset does not have 100 rows.") # Error for mismatched row count
 }
 
 # Check if the dataset has 13 columns
 if (ncol(analysis_data) == 13) {
-  message("Test Passed: The dataset has 13 columns.")
+  message("Test Passed: The dataset has 13 columns.") # Confirm column count matches
 } else {
-  stop("Test Failed: The dataset does not have 13 columns.")
+  stop("Test Failed: The dataset does not have 13 columns.") # Error for mismatched column count
 }
 
 # Check if all values in the 'X_id' column are unique
 if (n_distinct(analysis_data$X_id) == nrow(analysis_data)) {
-  message("Test Passed: All values in 'X_id' are unique.")
+  message("Test Passed: All values in 'X_id' are unique.") # Confirm uniqueness of IDs
 } else {
-  stop("Test Failed: The 'X_id' column contains duplicate values.")
[…]

### 2024-12-04 scripts/02-download_data.R

@@ -14,25 +14,26 @@ library(tidyverse)
 library(rstudioapi)
 
 #### Download data ####
-# get package
+# Retrieve metadata for the specified dataset using its unique package ID
 package <- show_package("1fc65d1e-7dae-4766-98dd-3b172e40a089")
-package
+package  # View metadata to confirm the package details
 
-# get all resources for this package
+# Get all resources associated with the package
 resources <- list_package_resources("1fc65d1e-7dae-4766-98dd-3b172e40a089")
 
-# identify datastore resources; by default, Toronto Open Data sets datastore resource format to CSV for non-geospatial and GeoJSON for geospatial resources
-# here I used csv format data
+# Filter resources to identify only those in CSV format, which are typically non-geospatial data
+# "datastore_resources" will hold only the CSV data resources from the package
 datastore_resources <- filter(resources, tolower(format) %in% c('csv'))
 
-
-# load the first datastore resource as a sample
-data <- filter(datastore_resources, row_number()==1) %>% get_resource()
-data
-
+# Load the first resource from the datastore as an example or sample dataset
+# "data" will contain the actual data from the first resource in CSV format
+data <- filter(datastore_resources, row_number() == 1) %>% get_resource()
+data  # View the loaded dataset to verify the structure and content
 
 #### Save data ####
-# change the_raw_data to whatever name you assigned when you downloaded it.
+# Set the working directory dynamically to the parent folder of the current script
 fd <- paste0(dirname(rstudioapi::getActiveDocumentContext()$path), "/../")  
 setwd(fd)
-write_csv(data, "./data/raw_data/raw_data.csv") 
+
+# Save the loaded dataset to a local directory in CSV format
+write_csv(data, "./data/raw_data/raw_data.csv")

### 2024-12-04 scripts/03-clean_data.R

@@ -13,29 +13,55 @@ library(tidyverse)
 library(rstudioapi)
 library(janitor)
 
-# Set data directory and read raw data
+#### Set data directory and read raw data ####
+# Dynamically set the working directory to the parent folder of the current script
 fd <- paste0(dirname(rstudioapi::getActiveDocumentContext()$path), "/../")  
 setwd(fd)
+
+# Read the raw data from the specified path
 raw_data <- read_csv("./data/raw_data/raw_data.csv")
+
+# Check the column names of the raw data to confirm structure and available variables
 names(raw_data)
+
 #### Clean data ####
-# Only variables of interest are retained. Please refer to the paper for the list of variables of interest and reasons.
-# Only records occurred after 2014 were retained
-# Any records with missing data were excluded
+# Data cleaning steps include:
+# 1. Retaining only the variables of interest for analysis.
+# 2. Filtering records to include only those that occurred after 2014.
+# 3. Removing any records with missing data.
 cleaned_data <-
   raw_data %>%
-  select(X_id,	OCC_DATE,	OCC_YEAR,	OCC_MONTH,	
-         OCC_DAY,	OCC_DOW,	OCC_HOUR,	DIVISION,	
-         PREMISES_TYPE,	UCR_CODE,	OFFENCE,	
-         LONG_WGS84,	LAT_WGS84) %>%
-  filter(OCC_YEAR > 2013,
-         OCC_YEAR != "None") %>%
+  select(
+    # Select only these columns from the raw data
+    X_id,  # Unique identifier for each record
+    OCC_DATE,  # Date of the occurrence
+    OCC_YEAR,  # Year of the occurrence
+    OCC_MONTH,  # Month of the occurrence
+    OCC_DAY,  # Day of the month when the occurrence happened
+    OCC_DOW,  # Day of the week of the occurrence
+    OCC_HOUR,  # Hour of the day for the occurrence
+    DIVISION,  # Police division where the occurrence was reported
+    PREMISES_TYPE,  # Type of premises where the occurrence took place
+    UCR_CODE,  # Uniform Crime Reporting code
+    OFFENCE,  # Type of offence reported
+    LONG_WGS84,  # Longitude coordinate of the occurrence
+    LAT_WGS84  # Latitude coordinate of the occurrence
+  ) %>%
+  filter(
+    # Retain only records from years after 2013
+    OCC_YEAR > 2013,
[…]

### 2024-12-04 scripts/04-test_analysis_data.R

@@ -19,8 +19,8 @@ setwd(fd)
 analysis_data <- read_csv("./data/analysis_data/analysis_data.csv")
 
 #### Test data ####
-
 # Check if the dataset has 13 columns
+# Ensures that the dataset structure includes exactly 13 columns as expected.
 if (ncol(analysis_data) == 13) {
   message("Test Passed: The dataset has 13 columns.")
 } else {
@@ -28,16 +28,17 @@ if (ncol(analysis_data) == 13) {
 }
 
 # Check if all values in the 'X_id' column are unique
+# Verifies that each record has a unique identifier to prevent duplication errors.
 if (n_distinct(analysis_data$X_id) == nrow(analysis_data)) {
   message("Test Passed: All values in 'X_id' are unique.")
 } else {
   stop("Test Failed: The 'X_id' column contains duplicate values.")
 }
 
-# Check if the 'DIVISION' column contains only valid Australian state names
+# Check if the 'DIVISION' column contains only valid Toronto police division names
+# Ensures that the division names in the dataset match the expected list of valid names.
 valid_DIVISION <- c("D11", "D12", "D13", "D14", "D22", "D23", "D31", "D32", "D33", 
                     "D41", "D42", "D43", "D51", "D52", "D53", "D54", "D55")
-
 if (all(analysis_data$DIVISION %in% valid_DIVISION)) {
   message("Test Passed: The 'DIVISION' column contains only valid Toronto police division names.")
 } else {
@@ -45,8 +46,8 @@ if (all(analysis_data$DIVISION %in% valid_DIVISION)) {
 }
 
 # Check if the 'OFFENCE' column contains only valid crime types of car theft cases
+# Validates that the offence types match the predefined categories for car theft cases.
 valid_OFFENCE <- c("Theft From Motor Vehicle Over", "Theft From Motor Vehicle Under")
-
 if (all(analysis_data$OFFENCE %in% valid_OFFENCE)) {
   message("Test Passed: The 'OFFENCE' column contains only valid crime types of car theft cases.")
 } else {
@@ -54,22 +55,24 @@ if (all(analysis_data$OFFENCE %in% valid_OFFENCE)) {
 }
 
 # Check if the 'PREMISES_TYPE' column contains only valid location types
+# Ensures that all entries in the 'PREMISES_TYPE' column match the expected set of valid premise types.
[…]

### 2024-12-04 scripts/05-exploratory_data_analysis.R

@@ -21,69 +21,73 @@ library(knitr)
 fd <- paste0(dirname(rstudioapi::getActiveDocumentContext()$path), "/../")  
 setwd(fd)
 analysis_data <- read_csv("./data/analysis_data/analysis_data.csv")
-
-#### data transformations####
+#### Data Transformations ####
+# Prepare a dataset with monthly counts of car thefts
 monthly_counts_all <- analysis_data %>%
   mutate(
     YEARMONTH = format(OCC_DATE, "%Y-%m")  # Extract year and month in "YYYY-MM" format
   ) %>%
   group_by(YEARMONTH) %>%
-  summarise(Crime_Count = n(), .groups = "drop") %>%
+  summarise(Crime_Count = n(), .groups = "drop") %>% # Count the number of crimes per month
   mutate(
-    time = as.numeric(as.Date(paste0(YEARMONTH, "-01"))),
-    action = ifelse(time >= as.numeric(as.Date("2024-01-01")),1,0))
-write_csv(monthly_counts_all,"./data/analysis_data/monthly_counts_all.csv")
+    time = as.numeric(as.Date(paste0(YEARMONTH, "-01"))), # Convert the monthly date to numeric for modeling
+    action = ifelse(time >= as.numeric(as.Date("2024-01-01")), 1, 0) # Create an indicator for pre- and post-action plan periods
+  )
+write_csv(monthly_counts_all, "./data/analysis_data/monthly_counts_all.csv") # Save the processed data
 
-####simple poisson regression####
+#### Simple Poisson Regression ####
+# Fit a Poisson regression model to examine the effect of time and action on crime counts
 poisson_model_all <- glm(
-  Crime_Count ~ time +action,
-  family = poisson(link = "log"),
+  Crime_Count ~ time + action, # Crime count as a function of time and action
+  family = poisson(link = "log"), # Use the Poisson family with a log link
   data = monthly_counts_all
 )
 
+# Display the model's coefficients and confidence intervals
 kable(summary(poisson_model_all)$coefficients)
 kable(confint(poisson_model_all))
 
-
-#####prediction####
-
+##### Prediction #####
+# Add predicted counts from the model to the dataset
 monthly_counts_all <- monthly_counts_all %>%
   mutate(
-    Predicted_Count = predict(poisson_model_all, type = "response")
+    Predicted_Count = predict(poisson_model_all, type = "response") # Predict crime counts based on the model
   )
 
-##### Plot actual vs. predicted counts####
-
[…]

### 2024-12-04 scripts/06-model_data.R

@@ -27,51 +27,50 @@ monthly_counts_all <- read_csv("./data/analysis_data/monthly_counts_all.csv")
 #read previous model#
 poisson_model_all <- readRDS(  file = "./models/first_model.rds"
 )
+#### Quasi-Likelihood Poisson Regression ####
+# Fit a Quasi-Poisson regression model to account for overdispersion
+poisson_model_all2 <- glm(
+  Crime_Count ~ time + action, # Model crime count as a function of time and action
+  family = quasipoisson(), # Use Quasi-Poisson family to handle overdispersion
+  data = monthly_counts_all # Use the processed dataset
+)
 
-####quasi-likelihood poisson regression####
-poisson_model_all2 <- glm(Crime_Count ~ time +action,
-                          family = quasipoisson(),
-                          data = monthly_counts_all)
+# Display the model's coefficients and confidence intervals
+kable(summary(poisson_model_all2)$coefficients) # Show regression coefficients
+kable(confint(poisson_model_all2)) # Show confidence intervals for the coefficients
 
-kable(summary(poisson_model_all2)$coefficients)
-kable(confint(poisson_model_all2))
-
-
-
-###prediction###
+### Prediction ###
+# Add predicted counts from the Quasi-Poisson model to the dataset
 monthly_counts_all <- monthly_counts_all %>%
   mutate(
-    Predicted_Count2 = predict(poisson_model_all2, type = "response")
+    Predicted_Count2 = predict(poisson_model_all2, type = "response") # Predict crime counts
   )
 
-
-
-##### Plot actual vs. predicted counts####
-
+##### Plot Actual vs. Predicted Counts ####
+# Visualize the actual and predicted crime counts over time using the Quasi-Poisson model
 ggplot(monthly_counts_all, aes(x = as.Date(time))) +
   geom_line(aes(y = Crime_Count, color = "Actual", linetype = "Actual"), size = 1) + # Actual counts
   geom_line(aes(y = Predicted_Count2, color = "Predicted", linetype = "Predicted"), size = 1) + # Predicted counts
-  scale_color_manual(values = c("Actual" = "blue", "Predicted" = "red")) + # Define line colors
-  scale_linetype_manual(values = c("Actual" = "solid", "Predicted" = "dashed")) + # Define line types
+  scale_color_manual(values = c("Actual" = "blue", "Predicted" = "red")) + # Set line colors
[…]

### 2024-12-04 paper/paper.qmd
correct typo

@@
 #| message: false
-
 library(tidyverse)
@@
 Vehicle theft has become a widespread public concern in Canada, leaving many citizens feeling helpless and frustrated.
-As car theft incidents continue to rise, police chiefs have made surprising recommendations, including advising residents to leave their car keys near the door to avoid violent confrontations during break-ins.
-This suggestion sparked public outrage, with many criticizing it online as absurd and unreasonable.
+As car theft incidents continue to rise, police chiefs have made surprising recommendations, including advising residents to leave their car keys near the door to avoid violent confrontations during break-ins @citynews2024.
+This suggestion sparked public outrage, with many criticizing it online as absurd and unreasonable @globalnews2024.
 The widespread public discourse escalated the issue into a national battle against crime.
@@
 The scale of vehicle theft in Canada is alarming, exposing government inefficiency in tackling the problem.
+In 2023, the number of motor vehicle thefts in Canada increased by 8.44% compared to the previous year, totaling 114,863 incidents @statista2024.
 The reaction from law enforcement has been equally unexpected, with officers treating theft as an inevitable occurrence and even providing rationales for its prevalence.
-While insurance companies offer quick resolutions for stolen vehicle claims, this comes at a cost.
+While insurance companies offer quick resolutions for stolen vehicle claims, this comes at a cost @ibc2024.
 Insurers are not charitable organizations, and frequent claims could lead to higher premiums or even the loss of insurance eligibility for some car owners.
@@
 The indifferent attitude of law enforcement, coupled with lenient legal consequences, has fueled an annual increase in car thefts, pushing Canada into a vicious cycle of buying and stealing vehicles.
-The legal responsibilities surrounding car theft and the evasive accountability of international import companies reflect the complexities of organized crime.
+The legal responsibilities surrounding car theft and the evasive accountability of international import companies reflect the complexities of organized crime @cbc2024.
 A large portion of stolen vehicles ends up in the Middle East and Africa, particularly Nigeria, where demand for high-quality second-hand cars is strong.
@@
 
-Under pressure from the public and media, the Canadian government introduced a series of policies in early 2024 as part of a “National Action Plan” to combat auto theft.
+Under pressure from the public and media, the Canadian government introduced a series of policies in early 2024 as part of a “National Action Plan” to combat auto theft @canadagov2024.
 The key components of the plan include:

### 2024-12-04 paper/paper.qmd
sketchupdate

@@
 #| message: false
+
 library(tidyverse)
@@
 # Render the table
-kable(variable_description, col.names = c("Column", "Description")) %>% kable_styling(full_width = FALSE) %>%  column_spec(1, width = "4cm") %>% column_spec(2, width = "8cm")
+kable(variable_description, col.names = c("Column", "Description"), format = "latex") %>% kable_styling(full_width = FALSE) %>%  column_spec(1, width = "4cm") %>% column_spec(2, width = "8cm")
 ```
@@
 dt <- read_csv(here("data","analysis_data","analysis_data.csv"))
-data_part1 <- kable(head(dt)[, 1:6],caption = "Sample of Clean Data: Part 1")
-data_part2 <- kable(head(dt)[, 7:10],caption = "Sample of Clean Data: Part 2")
-data_part3 <- kable(head(dt)[, 11:13],caption = "Sample of Clean Data: Part 3")
+data_part1 <- kable(head(dt)[, 1:6],caption = "Sample of Clean Data: Part 1", format = "latex")
+data_part2 <- kable(head(dt)[, 7:10],caption = "Sample of Clean Data: Part 2", format = "latex")
+data_part3 <- kable(head(dt)[, 11:13],caption = "Sample of Clean Data: Part 3", format = "latex")
 data_part1
@@
 )
-kable(summary(poisson_model_all)$coefficients)
+kable(summary(poisson_model_all)$coefficients, format = "latex")
 ```
@@
 dispersion_stat <- sum(residuals(poisson_model_all, type = "pearson")^2) / poisson_model_all$df.residual
-kable(data.frame(Stats="Dispersion Factor",Value=dispersion_stat))
+kable(data.frame(Stats="Dispersion Factor",Value=dispersion_stat), format = "latex")
 ```
@@
 
-kable(summary(poisson_model_all2)$coefficients)
+kable(summary(poisson_model_all2)$coefficients, format = "latex")
 ```
@@
 t2$model_type <- "Quasi-Poisson Model"
-kable(rbind(t1,t2))
+kable(rbind(t1,t2), format = "latex")
 ```
@@
 )
-kable(summary(poisson_model_PREMISES_TYPE)$coefficients)
+kable(summary(poisson_model_PREMISES_TYPE)$coefficients, format = "latex")
 ```
@@
 )
-kable(summary(poisson_model_OFFENCE)$coefficients)
+kable(summary(poisson_model_OFFENCE)$coefficients, format = "latex")
 ```
@@
 )
-kable(summary(poisson_model_DIVISION)$coefficients)
+kable(summary(poisson_model_DIVISION)$coefficients, format = "latex")
 ```

### 2024-12-15 paper/paper.qmd
LastCheck
Check before submission

@@
 ---
-title: "Canada’s Battle Against Car Theft: Progress Amid Persistent Challenges "
-subtitle: "Government measures and public vigilance show promise, but organized theft networks remain a concern."
+title: "Fight Against Car Theft: Progress and Challenges"
+subtitle: "A Data-Driven Review of the National Action Plan"
 author: 
@@
 date-format: long
-abstract: "Vehicle theft has become a significant public concern in Canada, highlighting both systemic challenges and promising governmental interventions.
-This study evaluates the effectiveness of the 2024 National Action Plan, which led to a 16.9% reduction in the mean monthly count of car thefts in Toronto.
-Spatial and temporal analyses reveal variations across police divisions and premise types, emphasizing the need for targeted strategies.
-Despite encouraging trends, organized theft networks and data limitations underscore the need for further research and adaptive policymaking."
+abstract: "Car theft remains a pressing public safety concern in Canada, prompting a national response through the 2024 National Action Plan.
+This study uses Toronto as a case example to evaluate the plan’s effectiveness, reporting a 16.93% reduction in the average monthly car theft count.
+Spatial and temporal analyses reveal notable disparities across areas covered by different police divisions and property types, highlighting the need for tailored local interventions.
+While the overall decline is encouraging, persistent challenges such as organized theft rings and data limitations underscore the necessity for continued research and adaptive policy development."
 include-before:
@@
     number-sections: true
+    mainfont: Arial
+    sansfont: Arial
+    fig-pos: 'H'
 bibliography: references.bib
-csl: apa6.csl
+csl: nature.csl
 editor_options: 
 chunk_output_type: console
+
 ---
@@
 
-Vehicle theft has become a widespread public concern in Canada, leaving many citizens feeling helpless and frustrated.
-As car theft incidents continue to rise, police chiefs have made surprising recommendations, including advising residents to leave their car keys near the door to avoid violent confrontations during break-ins @citynews2024.
-This suggestion sparked public outrage, with many criticizing it online as absurd and unreasonable @globalnews2024.
+Car theft has become a widespread public concern in Canada, leaving many citizens feeling helpless and frustrated.
+As car theft incidents continue to rise, police chiefs have made surprising recommendations, including advising residents to leave their car keys near the door to avoid violent confrontations during break-ins [@citynews2024].
+This suggestion sparked public outrage, with many criticizing it online as absurd and unreasonable [@globalnews2024].
 The widespread public discourse escalated the issue into a national battle against crime.
@@
 
-The scale of vehicle theft in Canada is alarming, exposing government inefficiency in tackling the problem.
+The scale of car theft in Canada is alarming, exposing government inefficiency in tackling the problem.
 In 2023, the number of motor vehicle thefts in Canada increased by 8.44% compared to the previous year, totaling 114,863 incidents @statista2024.
@@
 
-## The Vicious Cycle of Vehicle Theft
+## Global Demand
 
@@
 The legal responsibilities surrounding car theft and the evasive accountability of international import companies reflect the complexities of organized crime @cbc2024.
-A large portion of stolen vehicles ends up in the Middle East and Africa, particularly Nigeria, where demand for high-quality second-hand cars is strong.
+A large portion of stolen vehicles ends up in the Middle East and Africa, particularly Nigeria, where demand for high-quality used cars is strong.
 The COVID-19 pandemic and geopolitical conflicts have exacerbated supply shortages, causing car prices in the Middle East to surge and further incentivizing theft.
 
-## The Global Supply Chain of Stolen Vehicles
+## International Supply Chain
 
-The vehicle theft industry operates as an international supply chain, where cars of varying quality are allocated to different markets, creating a clear hierarchy.
-High-value vehicles are often shipped to affluent buyers, while lower-tier models serve other regions.
-The organized and systematic nature of these operations underscores the challenges faced by Canadian authorities in addressing this growing issue.
-Without comprehensive legal reforms and stricter enforcement measures, the country risks further entrenchment in this illicit cycle, with profound economic and social consequences.
+The car theft industry functions as a global supply chain, allocating vehicles of varying quality to different markets, forming a clear hierarchy.
+High-value cars are often shipped to wealthy buyers, while lower-tier models are directed to less affluent regions.
+Locally stolen vehicles are transported to seaports, illegally smuggled through shipping containers, and delivered to buyers worldwide.
+The organized and systematic nature of these operations highlights the significant challenges Canadian authorities face in combating this escalating issue.
+Without comprehensive legal reforms and stricter enforcement measures, the country risks deeper entrenchment in this illicit cycle, leading to severe social repercussions.
 
-## Public Response and Anti-Theft Measures
+## Public Response
 
-Faced with insufficient responses from the police, many vehicle owners have begun implementing their own anti-theft measures.
-However, professional thieves often view these devices as minor obstacles, and even robust anti-theft systems have limitations.
-High installation costs and vulnerabilities in extreme weather conditions further undermine their effectiveness.
-Some individuals attempt to protect their vehicles by parking in neighborhoods with higher police presence or near police stations.
-However, such measures do not guarantee safety, as evidenced by reports of police chiefs’ vehicles being stolen in broad daylight.
-These incidents highlight the audacity of thieves and have led to online ridicule of police authorities for failing to practice basic precautions.
+Faced with inadequate police responses, many vehicle owners have resorted to implementing their own anti-theft measures.
+Advanced security devices and traditional home-installed wheel locks are commonly used to deter theft.
+However, professional thieves often see these measures as minor obstacles, and even the most advanced systems have limitations.
+High installation costs and reduced effectiveness in extreme weather further compromise their reliability.
+Some owners attempt to safeguard their vehicles by parking in areas with a higher police presence or near police stations.
+Yet, such strategies offer no guarantee, as evidenced by reports of police chiefs' cars being stolen @belanger2023.
+These incidents expose the boldness of car thieves and have sparked online criticism of law enforcement for failing to uphold basic security practices.
 
@@
 
-- **Amendments to Criminal Laws**: Revising legislation to provide law enforcement and prosecutors with enhanced tools to tackle vehicle theft.
+- **Amendments to Criminal Laws**: Revising legislation to provide law enforcement and prosecutors with enhanced tools to tackle car theft.
 - **Changes to the Radiocommunication Act**: Tightening regulations on radio devices commonly used in car thefts.
@@
 - **Technology Deployment**: Leveraging advanced technologies such as scanning systems, data analytics, and GPS tracking to enhance inspection efficiency, particularly at ports and rail terminals.
-- **Specialized Training**: Offering targeted training programs for law enforcement officers to improve their capacity to investigate vehicle theft cases.
+- **Specialized Training**: Offering targeted training programs for law enforcement officers to improve their capacity to investigate car theft cases.
 - **Intergovernmental Task Force**: Creating a National Vehicle Theft Task Force to enhance coordination among government bodies.
@@
 
+## Role of Toronto
+
+Toronto plays a central role in Canada’s car theft crisis due to its large population, extensive vehicle ownership, and strategic location near major shipping ports.
+The city accounts for a significant percentage of national vehicle theft cases, driven by organized crime groups targeting both high-end and commonly used vehicles.
+Its proximity to international export routes facilitates the illegal shipping of stolen cars to overseas markets.
+Therefore, Toronto serves as an excellent case study for analyzing Canada’s car theft situation.
+
 ## Interim Result and Aim of the Paper
 
-Several months have passed since the plan's implementation.
-Open Data Toronto has published theft-from-motor-vehicle data for Toronto reported from January 2014 to September 2024.
+Several months have passed since the National Action Plan's implementation.
+Open Data Toronto has published theft-from-motor-vehicle data for Toronto reported from January 2014 to September 2024 @toronto_theft2024.
 To assess interim results, we analyzed this dataset to evaluate the effectiveness of the government's actions.
@@
 
-The primary estimand of this study is the monthly reduction in mean counts of car theft cases in Toronto following the implementation of the National Action Plan.
+The primary estimand of this study is the reduction in incidence of car theft cases in Toronto following the implementation of the National Action Plan.
 This is defined as the difference in the mean monthly car theft count before and after January 2024, when the plan starts.
-The estimand will be evaluated across all vehicle theft records reported in Toronto, stratified by key covariates.
+The estimand will be evaluated across all car theft records reported in Toronto, stratified by key covariates.
 This analysis assumes that the observed trends are not confounded by unrelated temporal factors.
@@
 Since 2014, the Toronto Police Service has been collecting auto theft data in digital format.
-The data is updated and maintained quarterly, with public access provided via Open Data Toronto.
+The data is updated and maintained quarterly, with public access provided via Open Data Toronto @toronto_theft2024.
 This dataset includes approximate locations of car theft incidents, facilitating the analysis of geographic trends over time.
@@
 
-The dataset used in this study was obtained from [Open Data Toronto](https://open.toronto.ca/dataset/deaths-of-shelter-residents/).
-The raw dataset comprises the following [key fields](@tbl-variable_description):
+The dataset used in this study was obtained from [Open Data Toronto](https://open.toronto.ca/dataset/theft-from-motor-vehicle/).
+The raw dataset comprises the key fields described in @tbl-variable_description:
 
@@
 Data cleaning steps are described in the [Appendix](#sec-appendix).
-A sample set of the cleaned data is presented in Table 2-4.
+A sample set of the cleaned data of first few variables is presented in @tbl-sample_data.
 Among all variables, Year, Total_Death, Male, and Female are coded as integers, Month is coded as abbreviations in character format, and Time is coded in date format.
@@
 
-```{r}
+```{r sampledata}
 #| echo: false
 #| warning: false
+#| label: tbl-sample_data
+#| tbl-cap: "Sample of Clean Data"
 
@@
 dt <- read_csv(here("data","analysis_data","analysis_data.csv"))
-data_part1 <- kable(head(dt)[, 1:6],caption = "Sample of Clean Data: Part 1", format = "latex")
-data_part2 <- kable(head(dt)[, 7:10],caption = "Sample of Clean Data: Part 2", format = "latex")
-data_part3 <- kable(head(dt)[, 11:13],caption = "Sample of Clean Data: Part 3", format = "latex")
+data_part1 <- kable(head(dt)[, 1:6])
 data_part1
-data_part2
-data_part3
 ```
@@
 
-The Poisson distribution is utilized to model the monthly count of car thefts, with a binary variable "action" serving as an indicator of whether the time of theft occurrence falls after the implementation of the national action plan.
+The Poisson distribution is utilized to model the monthly count of car thefts, with a binary variable "action" serving as an indicator of whether the time of theft occurrence falls after the implementation of the National Action Plan.
 The following are the formulas for the Poisson regression model:
@@
 $$
-where y_i is the count of car theft occurance.
+where $\lambda_i$ is the expectation of count of car theft occurrence, $\beta_0$ is the intercept, $\beta_i$ is the corresponding coefficient for a predictor.
 
@@
 This assessment included checks for overdispersion, identification of outliers, and analysis of residual distribution.
-Further details are provided in the [results section]{@sec-results}.
+Further details are provided in @sec-results.
 
@@
 
+@fig-Spatial-View illustrates the spatial distribution of car theft occurrences in September 2024 across police divisions.
+@fig-division-chart displays the overall count of car theft cases by division.
+Notably, Division D54 reported the lowest number of car theft cases, with Divisions D11 to D13 also showing relatively low counts, while Division D32 recorded the highest.
+
 ```{r}
@@
 #| warning: false
-#| fig-cap: "Spatial view of vehicle theft cases by division for September 2024"
+#| fig-cap: "Spatial view of car theft cases by division for September 2024"
 
@@
 ```
-
-@fig-Spatial-View illustrates the spatial distribution of car theft occurrences in September 2024 across police divisions.
-@fig-division-chart displays the overall count of car theft cases by division.
-Notably, Division D54 reported the lowest number of car theft cases, with Divisions D11 to D13 also showing relatively low counts, while Division D33 recorded the highest.
-Further analysis is needed, considering additional information about public safety conditions within these division areas.
 
@@
 
-@fig-premises-chart reveals that car theft cases occurred most frequently outside, followed by educational institutions, while transit locations reported the fewest cases.
+@fig-premises-chart reveals that car theft cases occurred most frequently outside, followed by houses, while transit locations reported the fewest cases.
 Across all premise types, two periods of increase were observed in 2019 and 2022-2023, with a decline noted after 2024.
@@
 
-@fig-offence-chart shows that most car theft cases were categorized as "Theft From Motor Vehicle Under," involving financial losses of less than $5,000.
-The number of severe cases remained relatively stable over the years, while lighter cases experienced two increases in 2019 and 2022, with declines observed in 2016 and after 2024.
+@fig-offence-chart shows that most car theft cases were categorized as "Theft From Motor Vehicle Under", involving financial losses of less than $5,000.
+The number of severe cases remained relatively stable over the years, while lighter cases experienced two increases in 2019 and 2022, with declines observed in 2016, 2021 and after 2024.
 
@@
 
-@fig-dow-chart indicates that car theft crimes were distributed relatively evenly across all days of the week, with a slight increase observed on Fridays.
+@fig-dow-chart indicates that car theft crimes were distributed evenly across all days of the week, with a slight increase observed on Fridays.
 
@@
 Interestingly, @fig-day-chart reveals that car theft cases occurred most frequently on the first day of the month and least often at the end of the month.
-The difference could be as high as 1,000 cases over a 10-year period, equivalent to approximately 30%.
+The difference could be as big as 1,000 cases over a 10-year period, equivalent to approximately 30%.
 
@@
 
-@tbl-Poisson indicates that the variable "action" had a significant effect on the mean count of car theft cases.
-Specifically, after the implementation of the national action plan, there was an approximate 1−exp(−0.185)=1−0.831=16.9% reduction in the mean count of car theft cases.
+@tbl-Poisson indicates that the variable "action", indicator of implimentation of National Action Plan, had a significant effect on the mean count of car theft cases.
+Specifically, after the implementation of the National Action Plan, there was an approximate $$1-exp(-0.1855)=1-0.8307=16.93\% $$ reduction in the mean count of car theft cases.
 This result highlights the potential effectiveness of the action plan in mitigating car theft, suggesting a notable decrease in incidents following its introduction.
@@
 #| tbl-cap: "Poisson regression results"
+#| tbl-pos: 'H'
 
@@
 )
-kable(summary(poisson_model_all)$coefficients, format = "latex")
+kable(round(summary(poisson_model_all)$coefficients,4), format = "latex")
 ```
@@
 Predictions from the model in @fig-Poisson-pred demonstrate a significant difference in the expected monthly car theft counts, with post-2024 counts being approximately 100 cases lower per month on average.
-This reduction suggests a measurable impact of national action plan starting in 2024.
+This reduction suggests a measurable impact of National Action Plan starting in 2024.
 
@@
 Diagnostic plots in @fig-Poisson-diagnosis show no major deflation in the QQ plot, suggesting a normal distribution of residuals.
-Leverage plots indicate only three significant outliers, corresponding to observations 5, 26, and 122.
+Leverage plots indicate only one influential point, corresponding to observation 122.
 
@@
 dispersion_stat <- sum(residuals(poisson_model_all, type = "pearson")^2) / poisson_model_all$df.residual
-kable(data.frame(Stats="Dispersion Factor",Value=dispersion_stat), format = "latex")
+kable(data.frame(Stats="Dispersion Factor",Value=round(dispersion_stat,4)), format = "latex")
 ```
@@
 #| fig-cap: "Poisson regression model diagnosis"
+#| fig.width: 8
+#| fig.height: 12
 
@@
 
-@tbl-Q-Poisson shows that the Quasi-Poisson model produced the same point estimate for the effect of "action," indicating a 16.9% reduction in the mean count of car theft cases following the implementation of the national action plan.
-However, the Quasi-Poisson model estimated a larger standard error, resulting in narrower confidence intervals (see @tbl-Q-Poisson), which may provide a more reliable and appropriate measure of precision.
+@tbl-Q-Poisson shows that the Quasi-Poisson model produced the same point estimate for the effect of "action", indicating a 16.93% reduction in the mean count of car theft cases following the implementation of the National Action Plan.
+However, the Quasi-Poisson model estimated a larger standard error, resulting in narrower confidence intervals (see @tbl-Q-P-CI), which provides a more reliable and appropriate measure of precision.
 
@@
 #| tbl-cap: "Quasi-likelihood Poisson regression results"
+#| tbl-pos: 'H'
 
@@
 
-kable(summary(poisson_model_all2)$coefficients, format = "latex")
+kable(round(summary(poisson_model_all2)$coefficients,4), format = "latex")
 ```
@@
 #| tbl-cap: "Comparison between Quasi-likelihood Poisson and Poisson estimated confidence intervals"
+#| tbl-pos: 'H'
 
@@
 
+@tbl-appendix-Q-P-PREMISES_TYPE, @tbl-appendix-Q-P-OFFENCE, and @tbl-appendix-Q-P-DIVISION in [Appendix] present point estimates from models incorporating different covariates.
+The results indicate that parking outside significantly increases the risk of car theft occurrences.
+Additionally, theft categorized as "Theft From Motor Vehicle Under" was associated with a higher number of car theft cases.
+Among the divisions, Division 32 exhibited the highest risk of car theft occurrence.
+
+# Discussion
+
+## Summary of Findings
+
+### Impact of the National Action Plan {#sec-first-point}
+
+The implementation of the National Action Plan in early 2024 has led to a measurable reduction in car theft in Toronto.
+Our analysis indicates a 16.93% decline in the average monthly count of car theft cases after the plan’s introduction, as shown by both Poisson and Quasi-Poisson regression models.
+Monthly crime counts fell by approximately 100 cases per month following the policy's enactment.
+Compared with the Poisson regression model, the Quasi-Poisson model produced the same point estimate but provided a more precise confidence interval by addressing overdispersion.
+These findings underscore the potential of government-led initiatives in reducing auto theft, though long-term monitoring is needed to evaluate sustained effects.
+
+### Spatial and Temporal Patterns of Car Theft {#sec-second-point}
+
+Spatial analysis revealed notable differences in car theft rates across Toronto's police divisions.
+Division 32 consistently experienced the highest theft rates due to its urban environment, heavy traffic, and role as a commercial hub, making it an attractive target for criminals.
+In contrast, Division 54 reported the lowest theft rates, likely due to its suburban setting, lower population density, and reduced vehicular activity, enabling more effective policing.
+Outdoor parking areas had the highest theft frequencies, reflecting elevated risks for vehicles parked outside.
+Temporally, thefts peaked in 2019 and 2022-2023, with a decline after 2024.
+A temporary reduction in early 2021 was likely caused by decreased mobility and public activity due to the COVID-19 pandemic.
+Notably, car thefts were most frequent at the beginning of the month and dropped significantly toward the month’s end, warranting further investigation into this pattern.
+
+### Effectiveness of the National Action Plan {#sec-third-point}
+
+Although the National Action Plan appears to have reduced overall car theft cases, its effectiveness likely varies by division and offense type.
+The higher occurrence of "Theft From Motor Vehicle Under" cases may be linked to less expensive vehicles with weaker anti-theft features, making them more susceptible to theft.
+Similarly, thefts involving residential properties and outdoor parking areas underscore the need for targeted security enhancements in these settings.
+These findings highlight the importance of tailoring crime prevention strategies to address distinct regional challenges.
+
+## Weaknesses and Limitations
+
+This study has several limitations.
+First, the analysis relied on reported cases, potentially underestimating the true extent of car theft due to unreported incidents.
+Second, the overdispersion identified in the Poisson model indicates variability that may not have been fully captured by the included predictors.
+While the Quasi-Poisson model improved standard error estimates, incorporating additional covariates such as socioeconomic factors or policing intensity could enhance model accuracy.
+Interaction effects were not examined in this study.
+Lastly, the relatively short timeframe since the National Action Plan’s implementation limits the ability to assess its long-term effects.
+
+## Recommendations and Future Work
+
+Based on the findings, we recommend that policymakers focus on areas with persistently high car theft rates, such as Division 32, to deploy targeted interventions.
+Efforts to improve public awareness of high-risk premise types, such as outdoor parking, could also reduce theft occurrences.
+Further research should explore the role of socioeconomic and environmental factors in shaping car theft trends, leveraging longer-term data to evaluate the sustained effectiveness of the National Action Plan.
+Additionally, integrating machine learning approaches could enhance predictive modeling and offer more granular insights into crime dynamics.
+
+\newpage
+
+\appendix
+
+# Appendix {#sec-appendix}
+
+## Data Cleaning Steps
+
+Raw data from Open Data Toronto were prepared, so we did not need to perform many cleaning steps. A few cleaning steps are as follows:
+
+- Only variables of interest were kept.
+- All records with all types of missing were excluded.
+- Only record indicating car theft cases occurrence after 2014 were kept.
+
+Details are described in a [script](https://github.com/vanessadyy/Car_Theft/blob/main/scripts/03-clean_data.R).
+\newpage
+
+## Additional Tables
+
 ```{r}
-#| label: tbl-Q-P-PREMISES_TYPE
+#| label: tbl-appendix-Q-P-PREMISES_TYPE
 #| echo: false
@@
 #| tbl-cap: "Quasi-likelihood Poisson regression with covariate of premise types"
+#| tbl-pos: 'H'
+
 monthly_counts_PREMISES_TYPE <- dt %>%
@@
 ```{r}
-#| label: tbl-Q-P-OFFENCE
+#| label: tbl-appendix-Q-P-OFFENCE
 #| echo: false
@@
 #| tbl-cap: "Quasi-likelihood Poisson regression with covariate of offence types"
+#| tbl-pos: 'H'
+
 monthly_counts_OFFENCE <- dt %>%
@@
 ```{r}
-#| label: tbl-Q-P-DIVISION
+#| label: tbl-appendix-Q-P-DIVISION
 #| echo: false
@@
 #| tbl-cap: "Quasi-likelihood Poisson regression with covariate of divisions"
+#| tbl-pos: 'H'
+
 monthly_counts_DIVISION <- dt %>%
@@
 
-@tbl-Q-P-PREMISES_TYPE, @tbl-Q-P-OFFENCE, and @tbl-Q-P-DIVISION present point estimates from models incorporating different covariates.
-The results indicate that parking outside significantly increases the risk of car theft occurrences.
-Additionally, theft categorized as "Theft From Motor Vehicle Under" was associated with a higher number of car theft cases.
-Among the divisions, Division 32 exhibited the highest risk of car theft occurrence.
-
-# Discussion
-
-## Impact of the National Action Plan {#sec-first-point}
-
-The results of this study indicate that the National Action Plan implemented in early 2024 has shown a measurable impact on reducing car theft in Toronto.
-Our analysis demonstrates a 16.9% reduction in the mean count of car theft cases after the action plan’s introduction, as evidenced by both Poisson and Quasi-Poisson regression models.
-This decline is further supported by the observed decrease in monthly crime counts after 2024, with an average reduction of 100 cases per month.
-These findings highlight the potential effectiveness of government-led initiatives in curbing auto theft, although additional time is needed to assess the plan’s long-term impact comprehensively.
-
-## Spatial and Temporal Patterns of Car Theft {#sec-second-point}
-
-The spatial analysis revealed substantial variation in car theft occurrences across police divisions.
-Division 32 consistently exhibited the highest risk of car theft, while Divisions 11, 12, 13, and 54 reported the lowest counts.
-These patterns may reflect underlying differences in socioeconomic conditions, policing strategies, or community vigilance within these divisions.
-Additionally, the analysis of premise types indicated that thefts occurred most frequently in outdoor locations, suggesting that vehicle owners parking outside face higher risks.
-Temporal trends revealed spikes in crime during certain years, notably 2019 and 2022–2023, followed by a decline in 2024.
-Interestingly, car thefts were most common at the beginning of the month, possibly linked to behavioral or logistical patterns among offenders.
-
-## Effectiveness of the National Action Plan {#sec-third-point}
-
-While the National Action Plan appears to have contributed to a decline in car theft cases, its effectiveness likely varies by division and offense type.
-The reduction in "Theft From Motor Vehicle Under" cases aligns with the broader trend but raises questions about whether certain measures disproportionately target less severe crimes.
-Furthermore, the use of enhanced technologies and intelligence sharing could be more impactful in divisions with historically higher crime rates, such as Division 32.
-These findings underscore the importance of tailoring interventions to the specific needs and challenges of different regions.
-
-## Weaknesses and Limitations
-
-This study has several limitations.
-First, the data used were limited to reported cases, potentially underestimating the true extent of car theft due to unreported incidents.
-Second, the overdispersion identified in the Poisson model suggests variability that may not have been fully captured by the predictors included in our analysis.
-Although the Quasi-Poisson model provided more appropriate standard error estimates, additional covariates, such as socioeconomic factors or policing intensity, could further improve model accuracy.
-Lastly, the relatively short timeframe since the National Action Plan’s implementation restricts our ability to evaluate its sustained effects.
-
-## Recommendations and Future Work
-
-Based on the findings, we recommend that policymakers focus on areas with persistently high car theft rates, such as Division 32, to deploy targeted interventions.
-Efforts to improve public awareness of high-risk premise types, such as outdoor parking, could also reduce theft occurrences.
-Further research should explore the role of socioeconomic and environmental factors in shaping car theft trends, leveraging longer-term data to evaluate the sustained effectiveness of the National Action Plan.
-Additionally, integrating machine learning approaches could enhance predictive modeling and offer more granular insights into crime dynamics.
-
-\newpage
-
-\appendix
-
-# Appendix {-}
-
-**Data Cleaning Steps**
-
-Raw data from Open Data Toronto were prepared, so we did not need to perform many cleaning steps. A few cleaning steps are as follows:
-
-- Only variables of interest were kept.
-- All records with all types of missing were excluded.
-- Only record indicating car theft cases occurrence after 2014 were kept.
-
 \newpage
